---
title: "Hands-on 8: Geographically Weighted Predictive Models"
author: Johsuan Huang
date: 2025-10-23
date-modified: "last-modified"
description: ""
format: html
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
  catche: true
---

# 1 Overview

Predictive modelling uses statistical learning or machine learning techniques to predict outcomes. By and large, the event one wants to predict is in the future. However, a set of known outcome and predictors (also known as variables) will be used to calibrate the predictive models.

Geospatial predictive modelling is conceptually rooted in the principle that the occurrences of events being modeled are limited in distribution. When geographically referenced data are used, occurrences of events are neither uniform nor random in distribution over space. There are geospatial factors (infrastructure, sociocultural, topographic, etc.) that constrain and influence where the locations of events occur. Geospatial predictive modeling attempts to describe those constraints and influences by spatially correlating occurrences of historical geospatial locations with environmental factors that represent those constraints and influences.

## 1.1 Learning outcome

In this in-class exercise, we will learn how to build predictive model by using geographical random forest method. By the end of this hands-on exercise, you will acquire the skills of:

-   preparing training and test data sets by using appropriate data sampling methods,

-   calibrating predictive models by using both geospatial statistical learning and machine learning methods,

-   comparing and selecting the best model for predicting the future outcome,

-   predicting the future outcomes by using the best model calibrated.

# 2 The Data

Four data sets will be used in this model building exercise, they are:

-   **Aspatial dataset**:

    -   HDB Resale data: a list of HDB resale transacted prices in Singapore from Jan 2017 onwards. It is in csv format which can be downloaded from Data.gov.sg.

-   **Geospatial dataset**:

    -   *MP14_SUBZONE_WEB_PL*: a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format. This data set was also downloaded from Data.gov.sg

-   **Locational factors with geographic coordinates**:

    -   Downloaded from **Data.gov.sg**.

        -   **Eldercare** data is a list of eldercare in Singapore. It is in shapefile format.

        -   **Hawker Centre** data is a list of hawker centres in Singapore. It is in geojson format.

        -   **Parks** data is a list of parks in Singapore. It is in geojson format.

        -   **Supermarket** data is a list of supermarkets in Singapore. It is in geojson format.

        -   **CHAS clinics** data is a list of CHAS clinics in Singapore. It is in geojson format.

        -   **Childcare service** data is a list of childcare services in Singapore. It is in geojson format.

        -   **Kindergartens** data is a list of kindergartens in Singapore. It is in geojson format.

    -   Downloaded from **Datamall.lta.gov.sg**.

        -   **MRT** data is a list of MRT/LRT stations in Singapore with the station names and codes. It is in shapefile format.

        -   **Bus stops** data is a list of bus stops in Singapore. It is in shapefile format.

-   **Locational factors without geographic coordinates**:

    -   Downloaded from **Data.gov.sg**.

        -   **Primary school** data is extracted from the list on General information of schools from data.gov portal. It is in csv format.

    -   Retrieved/Scraped from **other sources**

        -   **CBD** coordinates obtained from Google.

        -   **Shopping malls** data is a list of Shopping malls in Singapore obtained from [Wikipedia](https://en.wikipedia.org/wiki/List_of_shopping_malls_in_Singapore).

        -   **Good primary schools** is a list of primary schools that are ordered in ranking in terms of popularity and this can be found at [Local Salary Forum](https://www.salary.sg/2021/best-primary-schools-2021-by-popularity).

# 3 Getting Started

## 3.1 Install and load R packages

Before we get started, it is important for us to install the necessary R packages into R and launch these R packages into R environment.

The R packages needed for this exercise are as follows:

-   R package for building OLS and performing diagnostics tests
    -   [**olsrr**](https://olsrr.rsquaredacademy.com/)
-   R package for calibrating geographical weighted family of models
    -   [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/index.html)
-   R package for multivariate data visualisation and analysis
    -   [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)
-   Spatial data handling
    -   **sf**
-   Attribute data handling
    -   **tidyverse**, especially **readr**, **ggplot2** and **dplyr**
-   Choropleth mapping
    -   **tmap**

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse,knitr)
```

# 4 Geospatial Data Wrangling

## 4.1 Reading data file to rds

Reading the input data sets. It is in simple feature data frame.

```{r}
mdata <- read_rds("data/rawdata/mdata.rds")
kable(head(mdata))
```

## 4.2 Data Sampling

The entire data are split into training and test data sets with 65% and 35% respectively by using *initial_split()* of **rsample** package. rsample is one of the package of tigymodels.

```{r}
#| eval: false

set.seed(1234)
resale_split <- initial_split(mdata,prop = 0.65)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
#| eval: false

write_rds(train_data, "data/model/train_data.rds")
write_rds(test_data, "data/model/test_data.rds")
```

# 5 Computing Correlation Matrix

Before loading the predictors into a predictive model, it is always a good practice to use correlation matrix to examine if there is sign of multicolinearity.

```{r}
#| fig.width: 12
#| fig.height: 8
par(bg="#f6f6f6")

mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 1, 
                   method = "number", 
                   type = "upper")
```

The correlation matrix above shows that all the correlation values are below 0.8. Hence, there is **no sign of multicolinearity**.

# 6 Building a non-spatial multiple linear regression

```{r}

train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")
```

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

# 7 GWR predictive method

In this section, we will learn how to calibrate a model to predict HDB resale price by using geographically weighted regression method of [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/index.html) package.

## 7.1 Computing adaptive bandwidth

Firstly, `bw.gwr()` of **GWmodel** package will be used to determine the optimal bandwidth to be used. (CV method)

```{r}
#| eval: false
bw_adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
    WITHIN_1KM_PRISCH,
  data=train_data,
  approach="CV",
  kernel="gaussian",
  adaptive=TRUE,
  longlat=FALSE)

write_rds(bw_adaptive, "data/model/bw_adaptive.rds")
```

The result shows that 40 neighbour points will be the optimal bandwidth to be used if adaptive bandwidth is used for this data set.

## 7.2 Constructing the adaptive bandwidth gwr model

```{r}
bw_adaptive <- read_rds("data/model/bw_adaptive.rds")
```

Now, we can go ahead to calibrate the gwr-based hedonic pricing model by using adaptive bandwidth and Gaussian kernel as shown in the code chunk below.

```{r}
#| eval: false

gwr_adaptive <- gwr.basic(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
  data=train_data, # train model
  bw=bw_adaptive, #CV method
  kernel = 'gaussian', 
  adaptive=TRUE,
  longlat = FALSE)

write_rds(gwr_adaptive,"data/model/gwr_adaptive.rds")
```

## 7.3 Retrieve gwr output object

```{r}
gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")
```

The code below can be used to display the model output.

```{r}
gwr_adaptive
```

## 7.4 Computing adaptive bandwidth for the test data

```{r}
#| eval: false

gwr_bw_test_adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
    WITHIN_1KM_PRISCH,
  data=test_data,
  approach="CV",
  kernel="gaussian",
  adaptive=TRUE,
  longlat=FALSE)

write_rds(gwr_bw_test_adaptive, 
          "data/model/gwr_bw_test_adaptive.rds")
```

## 7.5 Computing predicted values of the test data

```{r}
gwr_bw_test_adaptive <- read_rds("data/model/gwr_bw_test_adaptive.rds")
```

```{r}
# Convert to SpatialPointsDataFrame
train_sp <- as(train_data, "Spatial")
test_sp <- as(test_data, "Spatial")
```

```{r}
#| eval: false

# Then predict using the fitted model
gwr_pred <- gwr.predict(
  formula = resale_price ~ floor_area_sqm + 
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER + 
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
    WITHIN_1KM_PRISCH,
  data = train_sp,
  predictdata = test_sp,
  bw = 40,
  kernel = 'gaussian',
  adaptive = TRUE,
  longlat = FALSE
)

write_rds(gwr_pred,"data/model/gwr_pred.rds")
```

::: callout-caution
The predict function didn't work and showed an error message:

"Error in gw.dist(fd.locat, focus = i, p, theta, longlat) : No regression point is fixed"
:::

# 8 Calibrating Random Forest Model

In this section, we will learn how to calibrate a model to predict HDB resale price by using random forest function of [**ranger**](https://cran.r-project.org/web/packages/ranger/index.html) package.

## 8.1 Preparing coordinates data

The code chunk below extract the x,y coordinates of the full, training and test data sets.

```{r}
#| eval: false
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

Before continue, we write all the output into rds for future used:

```{r}
#| eval: false

coords_train <- write_rds(
  coords_train, 
  "data/model/coords_train.rds" )
coords_test <- write_rds(
  coords_test, 
  "data/model/coords_test.rds" )
```

```{r}
coords_train <- read_rds("data/model/coords_train.rds")
coords_test <- read_rds("data/model/coords_test.rds")
```

## 8.2 Dropping geometry field

First, we will drop geometry column of the sf data.frame by using `st_drop_geometry()` of sf package.

```{r}
train_data_nogeom <- train_data %>% 
  st_drop_geometry()

test_data_nogeom <- test_data %>% 
  st_drop_geometry()
```

## 8.3 Calibrating a non-spatial random forest model

```{r}
#| eval: false

set.seed(1234)
rf <- ranger(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN + 
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS +  
    WITHIN_1KM_PRISCH,
  data=train_data_nogeom)

```

```{r}
#| eval: false

write_rds(rf, "data/model/rf.rds")
rf <- read_rds( "data/model/rf.rds")
```

```{r}
rf
```

# 9 Calibrating Geographical Weighted Random Forest Model

In this section, we will learn how to calibrate a model to predict HDB resale price by using `grf()` of [**SpatialML**](https://cran.r-project.org/web/packages/ranger/index.html) package.

## 9.1 Calibrating using training data

The code chunk below calibrate a geographic ranform forest model by using `grf()` of **SpatialML** package:

```{r}
#| eval: false

set.seed(1234)
gwRF_adaptive <- grf(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
  dframe=train_data_nogeom, 
  bw=40,
  kernel="adaptive",
  coords=coords_train)


```

```{r}
#| eval: false

write_rds(gwRF_adaptive, "data/model/gwRF_adaptive.rds")
```

```{r}
gwRF_adaptive <- read_rds("data/model/gwRF_adaptive.rds")
```

## 9.2 Predicting by using test data

### 9.2.1 Preparing the test data

```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```

### 9.2.2 Predicting with test data

```{r}
#| eval: false

gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data_nogeom, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)

GRF_pred <- write_rds(gwRF_pred, "data/model/GRF_pred.rds")
```

### 9.2.3 Converting the predicting output into a data frame

The output of the `predict.grf()` is a vector of predicted values. It is wiser to convert it into a data frame for further visualisation and analysis.

```{r}
GRF_pred <- read_rds("data/model/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

In the code chunk below, `cbind()` is used to append the predicted values onto the test_data:

```{r}
test_data_p <- cbind(
  test_data, GRF_pred_df) %>%
  select(resale_price, GRF_pred)
```

```{r}
write_rds(test_data_p, "data/model/test_data_p.rds")
```

## 9.3 Calculating Root Mean Square Error

The root mean square error (RMSE) allows us to measure how far predicted values are from observed values in a regression analysis. In the code chunk below, rmse() of Metrics package is used to compute the RMSE.

```{r}
rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

## 9.4 Visualising the predicted values

Alternatively, scatterplot can be used to visualise the actual resale price and the predicted resale price by using the code chunk below.

```{r}
ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()+
  geom_smooth(method = lm)+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())
```

# 10 Reference

-   Kam, T.S. (2025). [Geographically Weighted Predictive Models](https://r4gdsa.netlify.app/chap14.html)
