---
title: "In-class 8: Take-home 3 kick-strat"
author: Johsuan Huang
date: 2025-10-23
date-modified: "last-modified"
description: ""
format: html
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
  catche: true
---

# 1 Import the packages

-   jsonlite: JSON parser and generator
-   progess: Configurable ***Progress*** bars, they may include percentage, elapsed time, and/or the estimated completion time

```{r}
pacman::p_load(httr,jsonlite,sf, spdep, progress, tmap, tidyverse,knitr, spdep,GWmodel, rsample, SpatialML,Metrics,kableExtra, randomForestExplainer,performance)
```

# 2 Geocode HDB resale data

```{r}
#| eval: false

HDBresale <- read_csv("data/rawdata/ResaleflatpricesbasedonregistrationdatefromJan2017onwards.csv")%>%
  filter(flat_type=="4 ROOM")%>%
  filter(month >="2025-09" & month < "2025-10")
```

```{r}
#| eval: false

HDBresale$street_name <- gsub("ST\\.","SAINT",HDBresale$street_name) 
```

```{r}
#| eval: false

geocode <- function(block, streetname){
  base_url <- "https://www.onemap.gov.sg/api/common/elastic/search"
  address <- paste(block,streetname,sep =" ")
  query <- list("searchVal" = address,
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" ="1")
  res <- GET(base_url, query =query)
  restext <- content(res, as="text")
  
  output <- fromJSON(restext) %>%
    as.data.frame %>%
    select(results.LATITUDE, results.LONGITUDE)
  
  return(output)
}
```

```{r}
#| eval: false

HDBresale$LATITUDE <- 0
HDBresale$LONGTITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i,4],HDBresale[i,5])
  
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGTITUDE[i] <- temp_output$results.LONGITUDE
}

```

```{r}
#| eval: false

write_rds(HDBresale,"data/HDBresale.rds")
```

```{r}
HDBresale <- read_rds("data/HDBresale.rds")
```

# 3 Prepare Data

```{r}

HDBresale_sf <- st_as_sf(HDBresale, 
                         coords = c("LONGTITUDE", "LATITUDE"),
                         crs = 4326) 
```

```{r}

HDBresale_sf <- HDBresale_sf %>%
  st_transform(3414)
```

```{r}
# only need to count the # of preschool
preschool <- st_read("data/rawdata/PreSchoolsLocation.kml")%>%
  st_transform(crs=3414)
```

```{r}
within_350m <- st_is_within_distance(HDBresale_sf, preschool, dist=350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(preschool_350m = map_int(within_350m, length))
```

# 4 Calibrate Predictive Model

## 4.1 Import Raw Data

```{r}
mdata <- read_rds("data/rawdata/mdata.rds")
```

## 4.2 Data Sampling

Since predictive model like random forest is computational intensive, only 10% data will be randomly selected for predictive modelling:

```{r}
set.seed(1234)
HDB_sample <- mdata %>% sample_n(1500)
```

## 4.3 Check of overlapping point

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(.,.)) >1 ) 
# The `.` represents the data being piped in
```

```{r}
kable(head(overlapping_points%>%
        filter(overlap == "TRUE")))
```

To avoid model crashes, we need to jitter overlapped points:

```{r}
HDB_sample <- HDB_sample %>%
  st_jitter(amount = 1) #jitter 1 meter apart
```

Let's check if there are any overlapping points again:

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(.,.)) >1 )

kable(head(overlapping_points%>%
        filter(overlap == "TRUE")))
```

```{r}
set.seed(1234)
resale_split <- initial_split(HDB_sample,
                              prop = 6.67/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

It's not necessary to build mlr if we are going to build models using GWmodel.

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

```{r}
check_collinearity(price_mlr)
```

```{r}
#| fig.width: 12
#| fig.height: 8
mlr.vif <- check_collinearity(price_mlr)
par(bg="#f6f6f6")
plot(mlr.vif) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 4.4 Import GWR model from Hands-on Ex08

```{r}
gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")
gwr_adaptive
```

The R-squared of global model (using full dataset) is 0.73 while the R-squared of localized geospatial model (using neighbors' dataset only) is 0.97, suggesting including geospatial info is important for predictive model.

For explanatory model, we need to look at p-value, R-squared, AIC for comparison between global and local models.

we can plot the significance of the variables using resource below:

<https://www.jstatsoft.org/article/view/v063i17>

```{r}
gwr_adaptive
```
