---
title: "In-class 9: Modelling Geographical Accessibility Using Hanson Method"
author: Johsuan Huang
date: 2025-11-08
date-modified: "last-modified"
description: ""
format: html
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
  catche: true
---

# 1 Overview

In this hands-on exercise, we will gain hands-on experience on how to model geographical accessibility by using R’s geospatial analysis packages

## 1.1 Learning outcome

By the end of this hands-on exercise, we will be able:

-   to import GIS polygon data into R and save them as simple feature data frame by using appropriate functions of sf package of R;
-   to import aspatial data into R and save them as simple feature data frame by using appropriate functions of sf package of R;
-   to computer accessibility measure by using Hansen’s potential model and Spatial Accessibility Measure (SAM); and
-   to visualise the accessibility measures by using tmap and ggplot2 packages.

# 2 The Data

Four data sets will be used in this model building exercise, they are:

-   `MP14_SUBZONE_NO_SEA_PL`: URA Master Plan 2014 subzone boundary GIS data. This data set is downloaded from data.gov.sg.

-   `hexagons`: A 250m radius hexagons GIS data. This data set was created by using [*st_make_grid()*](https://r-spatial.github.io/sf/reference/st_make_grid.html) of sf package. It is in ESRI shapefile format.

-   `ELDERCARE`: GIS data showing location of eldercare service. [This data](https://data.gov.sg/dataset/eldercare-services) is downloaded from data.gov.sg. There are two versions. One in ESRI shapefile format. The other one in Google kml file format. For the purpose of this hands-on exercise, ESRI shapefile format is provided.

-   `OD_Matrix`: a distance matrix in csv format. There are six fields in the data file. They are:

    -   `origin_id`: the unique id values of the origin (i.e. `fid` of hexagon data set.),

    -   `destination_id`: the unique id values of the destination (i.e. `fid` of `ELDERCARE` data set.),

    -   `entry_cost`: the perpendicular distance between the origins and the nearest road),

    -   `network_cost`: the actual network distance from the origin and destination,

    -   `exit_cost`: the perpendicular distance between the destination and the nearest road), and

    -   `total_cost`: the summation of `entry_cost`, `network_cost`and `exit_cost`.

All the values of the cost related fields are in **metres**

# 3 Getting Started

## 3.1 Install and load R packages

Before we get started, it is important for us to install the necessary R packages into R and launch these R packages into R environment.

The R packages needed for this exercise are as follows:

-   Spatial data handling
    -   sf
-   Modelling geographical accessibility
    -   spatialAcc
-   Attribute data handling
    -   tidyverse, especially readr and dplyr
-   thematic mapping
    -   tmap
-   Staistical graphic
    -   ggplot2
-   Statistical analysis
    -   ggstatsplot

The code chunk below installs and launches these R packages into RStudio environment:

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,tidyverse,knitr)
```

Notice that with tidyverse, we do not have to install readr, dplyr and ggplots packages separately. In fact, tidyverse also installs other R packages such as tidyr, stringr, forcats, tibble, purrr and magrittr.

# 4 Geospatial Data Wrangling

## 4.1 Importing geospatial data

Three geospatial data will be imported from the *data/geospatial* sub-folder. They are MP14_SUBZONE_NO_SEA_PL, hexagons and ELDERCARE.

The code chunk below is used to import these three data sets shapefile by using *st_read()* of **sf** packages.

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
```

```{r}
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
```

```{r}
head(kable(hexagons))
```

```{r}
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 
```

```{r}
head(kable(eldercare))
```

## 4.2 Updating CRS Information

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

After transforming the projection metadata, you can verify the projection of the newly transformed *mpsz_svy21* by using *st_crs()* of sf package.

The code chunk below will be used to verify the newly transformed *mpsz_svy21*.

```{r}
st_crs(mpsz)
```

Notice that the EPSG: is indicated as *3414* now.

## 4.3 Cleaning and updating attribute fields of the geospatial data

There are many redundant fields in the data tables of both `eldercare`and `hexagons`. The code chunks below will be used to exclude those redundant fields. At the same time, a new field called `demand` and a new field called `capacity` will be added into the data table of `hexagons` and `eldercare` sf data frame respectively. Both fields are derive using *mutate()* of **dplyr** package.

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

```{r}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

Notice that for the purpose of this hands-on exercise, a constant value of 100 is used. In practice, actual demand of the hexagon and capacity of the eldercare centre should be used.

# 5 Apsaital Data Handling and Wrangling

## 5.1 Importing Distance Matrix

The code chunk below uses *read_cvs()* of **readr** package to import `OD_Matrix.csv` into RStudio. The imported object is a tibble data.frame called `ODMatrix`.

```{r}
#skip=0 means do not skip any lines at the beginning of the file.
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0) 
```

## 5.2 Tidying distance matrix

The imported ODMatrix organised the distance matrix columnwise. On the other hands, most of the modelling packages in R is expecting a matrix look similar to the figure below.

On the other hands, most of the modelling packages in R is expecting a matrix look similar to the figure below.

The rows represent **origins** (i.e. also know as **from** field) and the columns represent **destination** (i.e. also known as **to** field.)

The code chunk below uses *spread()* of **tidyr** package is used to transform the O-D matrix from a thin format into a fat format.

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  pivot_wider(
    names_from = destination_id,
    values_from = total_cost
  ) %>%
  select(-origin_id)
```

Note: Since tidyr version 1.0 a new function called [*pivot_wider()*](https://tidyr.tidyverse.org/reference/pivot_wider.html) is introduce. You should use *pivot_wider()* instead of *spread()*

Currently, the distance is measured in metre because SVY21 projected coordinate system is used. The code chunk below will be used to convert the unit f measurement from metre to kilometre.

```{r}
distmat_km <- as.matrix(distmat/1000)
```

# 6 Modelling and Visualising Accessibility using Hansen Method

## 6.1 Computing Hansen’s accessibility

Now, we ready to compute Hansen’s accessibility by using *ac()* of [**SpatialAcc**](https://cran.r-project.org/web/packages/SpatialAcc/index.html) package. Before getting started, you are encourage to read the [arguments](https://cran.r-project.org/web/packages/SpatialAcc/SpatialAcc.pdf) of the function at least once in order to ensure that the required inputs are available.

The code chunk below calculates Hansen’s accessibility using *ac()* of **SpatialAcc** and *data.frame()* is used to save the output in a data frame called `acc_Handsen`.

observation: what changed between pow 0.5\~2.5

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 2.5, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "acc_Hansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen_2.5 <- bind_cols(hexagons, acc_Hansen)

```

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 2, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "acc_Hansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen_2 <- bind_cols(hexagons, acc_Hansen)
```

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 1.5, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "acc_Hansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen_1.5 <- bind_cols(hexagons, acc_Hansen)
```

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 1, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "acc_Hansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen_1 <- bind_cols(hexagons, acc_Hansen)
```

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 0.5, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "acc_Hansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen_0.5 <- bind_cols(hexagons, acc_Hansen)
```

::: callout-note
## Hansen’s Accessibility

Accessibility = sum of opportunities weighted by **distance decay**. Closer and larger facilities contribute more to accessibility than smaller or distant ones. It only considers supply.
:::

## 6.2 Visualising Hansen’s accessibility

### 6.2.1 Extracting map extend

Firstly, we will extract the extend of `hexagons` simple feature data frame by using *st_bbox()* of **sf** package.

```{r}
mapex <- st_bbox(hexagons)
```

The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.

```{r}
#| fig.width: 12
#| fig.height: 20
pow_0.5 <- tm_shape(hexagon_Hansen_0.5,
         bbox = mapex) + 
  tm_polygons(fill = "acc_Hansen", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "pow=0.5")) +
tm_shape(eldercare) +
  tm_dots(size = 0.3,
          fill = "red",
          col = "black") +
  tm_layout(frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.1,col="black") + 
  tm_layout(
    bg.color = "#f6f6f6",
    outer.bg.color = "#f6f6f6",
    frame = FALSE)

pow_1 <- tm_shape(hexagon_Hansen_1,
         bbox = mapex) + 
  tm_polygons(fill = "acc_Hansen", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "pow=1")) +
tm_shape(eldercare) +
  tm_dots(size = 0.3,
          fill = "red",
          col = "black")+
  tm_layout(frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.1,col="black") + 
  tm_layout(
    bg.color = "#f6f6f6",
    outer.bg.color = "#f6f6f6",
    frame = FALSE)

pow_1.5 <- tm_shape(hexagon_Hansen_1.5,
         bbox = mapex) + 
  tm_polygons(fill = "acc_Hansen", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "pow=1.5")) +
tm_shape(eldercare) +
  tm_dots(size = 0.3,
          fill = "red",
          col = "black")+
  tm_layout(frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.1,col="black") + 
  tm_layout(
    bg.color = "#f6f6f6",
    outer.bg.color = "#f6f6f6",
    frame = FALSE)

pow_2 <- tm_shape(hexagon_Hansen_2,
         bbox = mapex) + 
  tm_polygons(fill = "acc_Hansen", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "pow=2")) +
tm_shape(eldercare) +
  tm_dots(size = 0.3,
          fill = "red",
          col = "black")+
  tm_layout(frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.1,col="black") + 
  tm_layout(
    bg.color = "#f6f6f6",
    outer.bg.color = "#f6f6f6",
    frame = FALSE)

pow_2.5 <- tm_shape(hexagon_Hansen_2.5,
         bbox = mapex) + 
  tm_polygons(fill = "acc_Hansen", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "pow=2.5")) +
tm_shape(eldercare) +
  tm_dots(size = 0.3,
          fill = "red",
          col = "black")+
  tm_layout(frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.1,col="black") + 
  tm_layout(
    bg.color = "#f6f6f6",
    outer.bg.color = "#f6f6f6",
    frame = FALSE)

tmap_arrange(pow_0.5,pow_1,pow_1.5,ncol = 1)
```

```{r}
#| fig.width: 12
#| fig.height: 12
tmap_arrange(pow_2,pow_2.5)
```

As power increases, the overall accessibility scores decrease sharply. Even the best-served areas (in darker blue) see declining accessibility scores, because distant facilities contribute almost nothing.

With pow=0.5, eldercare facilities have a wider influence zone, meaning even facilities farther away still contribute significantly to a location's accessibility score. However, as power increase, the distance acts as a stronger barrier. Only areas very close to eldercare facilities show high accessibility. The penalty for distance is more severe, creating sharper spatial inequalities.

This suggests that the choice of power parameter is critical and context-dependent, it should be carefully calibrated based on empirical evidence of actual travel behavior rather than arbitrary selection.
