---
title: "Hands-on 7: Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method Techniques"
author: Johsuan Huang
date: 2025-10-15
date-modified: "last-modified"
description: ""
format: html
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
  catche: true
---

# 1 Overview

**Geographically weighted regression (GWR)** is a spatial statistical technique that takes non-stationary variables into consideration (e.g., climate; demographic factors; physical environment characteristics) and models the local relationships between these independent variables and an outcome of interest (also known as dependent variable). In this hands-on exercise, you will learn how to build [hedonic pricing](https://www.investopedia.com/terms/h/hedonicpricing.asp) models by using GWR methods. The dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

# 2 The Data

Two data sets will be used in this model building exercise, they are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)
-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

# 3 Getting Started

## 3.1 Install and load R packages

Before we get started, it is important for us to install the necessary R packages into R and launch these R packages into R environment.

The R packages needed for this exercise are as follows:

-   R package for building OLS and performing diagnostics tests
    -   [**olsrr**](https://olsrr.rsquaredacademy.com/)
-   R package for calibrating geographical weighted family of models
    -   [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/index.html)
-   R package for multivariate data visualisation and analysis
    -   [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)
-   Spatial data handling
    -   **sf**
-   Attribute data handling
    -   **tidyverse**, especially **readr**, **ggplot2** and **dplyr**
-   Choropleth mapping
    -   **tmap**

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, 
               sf, sfdep, GWmodel, tmap, 
               tidyverse, gtsummary,
               performance, RColorBrewer,see)
```

# 4 Geospatial Data Wrangling

## 4.1 Import shapefile

The geospatial data used in this hands-on exercise is called MP14_SUBZONE_WEB_PL. It is in ESRI shapefile format. The shapefile consists of URA Master Plan 2014’s planning subzone boundaries. Polygon features are used to represent these geographic boundaries.

The code chunk below is used to import *MP_SUBZONE_WEB_PL* shapefile by using `st_read()`of **sf** packages:

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
               layer = "MP14_SUBZONE_WEB_PL") %>%
  st_transform(3414)
st_crs(mpsz)
```

After importing the shape file, let's check the extent of mpsz and plot the graph:

```{r}
st_bbox(mpsz)
```

The print above reports the extent of mpsz layer by its lower and upper limits.

```{r}
par(bg="#f6f6f6")
plot(mpsz)
```

# 5 Aspatial Data Wrangling

## 5.1 Import the aspatial data

The *condo_resale_2015* is in csv file format. The codes chunk below uses `read_csv()` function of **readr** package to import *condo_resale_2015* into R as a tibble data frame called *condo_resale*.

```{r}
condo_resale <- read_csv("data/aspatial/Condo_resale_2015.csv")
```

There are a total of 23 fields and 1436 observation in the tibble data.frame.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE)
head(condo_resale$LATITUDE)
```

The print above reveals that the values of *LONGITITUDE* and *LATITUDE* fields are in decimal degree. Most probably wgs84 geographic coordinate system is used.

Next, `summary()` of base R is used to display the summary statistics of *cond_resale* tibble data frame.

```{r}
summary(condo_resale)
```

## 5.2 Convert aspatial data frame into a sf object

Currently, the *condo_resale* tibble data frame is aspatial. We will convert it to a **sf** object. The code chunk below converts condo_resale data frame into a simple feature data frame by using `st_as_sf()` of **sf** packages.

Notice that `st_transform()` of **sf** package is used to convert the coordinates from wgs84 (i.e. crs:4326) to svy21 (i.e. crs=3414).

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>% #wgs84
  st_transform(crs=3414) #svy21
```

Next, `head()` is used to list the content of *condo_resale.sf* object.

```{r}
head(condo_resale.sf)
```

# 6 Exploratory Data Analysis

## 6.1 EDA using statistical graphics

We can plot the distribution of *SELLING_PRICE* by using appropriate Exploratory Data Analysis (EDA) as shown in the code chunk below.

```{r}
ggplot(data=condo_resale.sf, 
                     aes(x=SELLING_PRICE)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  labs(title = "Histogram of Condo Selling Price",
       x = "", y = "Count") +
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())
```

The figure above reveals a right skewed distribution. This means that more condominium units were transacted at relative lower prices.

Statistically, the skewed dsitribution can be normalised by using log transformation. The code chunk below is used to derive a new variable called *LOG_SELLING_PRICE* by using a log transformation on the variable *SELLING_PRICE*. It is performed using `mutate()` of **dplyr**package.

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

```{r}
ggplot(data=condo_resale.sf, aes(x=LOG_SELLING_PRICE)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  labs(title = "Log of Condo Selling Price",
       x = "", y = "Count") +
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())
```

Notice that the distribution is relatively less skewed after the transformation.

## 6.2 Multiple Histogram Plots distribution of variables

In this section, we will learn how to draw a small multiple histograms (also known as trellis plot) by using `ggarrange()` of [**ggpubr**](https://cran.r-project.org/web/packages/ggpubr/index.html) package.

The code chunk below is used to create 12 histograms. Then, `ggarrange()` is used to organised these histogram into a 3 columns by 4 rows small multiple plot.

```{r}
#| fig.width: 12
#| fig.height: 10

AREA_SQM <- ggplot(data=condo_resale.sf, 
                   aes(x= AREA_SQM)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

AGE <- ggplot(data=condo_resale.sf, 
              aes(x= AGE)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_CBD <- ggplot(data=condo_resale.sf, 
                   aes(x= PROX_CBD)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, 
                         aes(x= PROX_CHILDCARE)) + 
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, 
                           aes(x= PROX_ELDERLYCARE)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= PROX_URA_GROWTH_AREA)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, 
                             aes(x= PROX_HAWKER_MARKET)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, 
                            aes(x= PROX_KINDERGARTEN)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_MRT <- ggplot(data=condo_resale.sf, 
                   aes(x= PROX_MRT)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_PARK <- ggplot(data=condo_resale.sf, 
                    aes(x= PROX_PARK)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                           aes(x= PROX_PRIMARY_SCH)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= PROX_TOP_PRIMARY_SCH)) +
  geom_histogram(bins=20, 
                 color="grey30", 
                 fill="#93A4CC")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, 
          PROX_ELDERLYCARE, PROX_URA_GROWTH_AREA,
          PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

## 6.3 Drawing Statistical Point Map

Lastly, we want to reveal the geospatial distribution condominium resale prices in Singapore. The map will be prepared by using **tmap** package.

First, we will turn on the interactive mode of tmap by using the code chunk below.

```{r}
tmap_mode("plot")
```

Next, the code chunks below is used to create an interactive point symbol map:

```{r}
tm_shape(mpsz)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(fill = "SELLING_PRICE",
          fill_alpha = 0.6,
          size = 0.5,
          fill.scale = tm_scale_intervals(
            style="quantile")) +
  tm_view(set_zoom_limits = c(11,14))
```

Notice that [`tm_dots()`](https://www.rdocumentation.org/packages/tmap/versions/2.2/topics/tm_symbols) is used instead of `tm_bubbles()`.

`set_zoom_limits` argument of `tm_view()` sets the minimum and maximum zoom level to 11 and 14 respectively.

Before moving on to the next section, the code below will be used to turn R display into `plot`mode.

```{r}
tmap_mode("plot")
```

# 7 Hedonic Pricing Modelling in R

In this section, we will learn how to building hedonic pricing models for condominium resale units using [`lm()`](https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/lm) of R base.

## 7.1 Simple Linear Regression Method

First, we will build a simple linear regression model by using *SELLING_PRICE* as the dependent variable and *AREA_SQM* as the independent variable:

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, 
                data = condo_resale.sf)
```

`lm()` returns an object of class “lm” or for multiple responses of class c(“mlm”, “lm”).

The functions `summary()` and `anova()` can be used to obtain and print a summary and analysis of variance table of the results. The generic accessor functions coefficients, effects, fitted.values and residuals extract various useful features of the value returned by `lm`.

```{r}
summary(condo.slr)
```

The output report reveals that the SELLING_PRICE can be explained by using the formula:

```         
      *y = -258121.1 + 14719x1*
```

The R-squared of 0.4518 reveals that the simple regression model built is able to explain about 45% of the resale prices.

Since p-value is much smaller than 0.0001, we will reject the null hypothesis that mean is a good estimator of SELLING_PRICE. This will allow us to infer that simple linear regression model above is a good estimator of *SELLING_PRICE*.

The **Coefficients:** section of the report reveals that the p-values of both the estimates of the Intercept and ARA_SQM are smaller than 0.001. In view of this, the null hypothesis of the B0 and B1 are equal to 0 will be rejected. As a results, we will be able to infer that the **B0 and B1 are good parameter estimates**.

To visualise the best fit curve on a scatterplot, we can incorporate `lm()` as a method function in ggplot’s geometry as shown in the code chunk below:

```{r}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())
```

Figure above reveals that there are a few statistical outliers with relatively high selling prices

## 7.2 Multiple Linear Regression Method

### 7.2.1 Visualise the correlation between independent variables

Before building a multiple regression model, it is important to ensure that the indepdent variables used are not highly correlated to each other. If these highly correlated independent variables are used in building a regression model by mistake, the quality of the model will be compromised. This phenomenon is known as **multicollinearity** in statistics.

Correlation matrix is commonly used to visualise the relationships between the independent variables. Beside the `pairs()` of R, there are many packages support the display of a correlation matrix. In this section, the [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html) package will be used.

The code chunk below is used to plot a scatterplot matrix of the relationship between the independent variables in *condo_resale* data.frame.

```{r}
#| fig.width: 12
#| fig.height: 10

par(bg="#f6f6f6")
corrplot(cor(condo_resale[, 5:23]), 
         diag = FALSE, 
         order = "AOE",
         tl.pos = "td", 
         tl.cex = 0.5, 
         method = "number", 
         type = "upper")
```

Matrix reorder is very important for mining the hiden structure and patter in the matrix. There are four methods in corrplot (parameter order), named “AOE”, “FPC”, “hclust”, “alphabet”. In the code chunk above, AOE order is used. It orders the variables by using the *angular order of the eigenvectors* method suggested by [Michael Friendly](https://www.datavis.ca/papers/corrgram.pdf).

From the scatterplot matrix, it is clear that ***Freehold*** is highly correlated to ***LEASE_99YEAR (-0.84)***. In view of this, it is wiser to only include either one of them in the subsequent model building. As a result, ***LEASE_99YEAR*** is excluded in the subsequent model building.

## 7.3 Build a hedonic pricing model using multiple linear regression method

The code chunk below using `lm()` to calibrate the multiple linear regression model:

```{r}
condo.mlr <- lm(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
    PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET   +
    PROX_KINDERGARTEN   + PROX_MRT  + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_TOP_PRIMARY_SCH +
    PROX_SHOPPING_MALL  + PROX_SUPERMARKET + 
    PROX_BUS_STOP   + NO_Of_UNITS + 
    FAMILY_FRIENDLY + FREEHOLD, 
  data=condo_resale.sf)
summary(condo.mlr) #print the model output
```

## 7.4 Revise the model

With reference to the report above, it is clear that not all the independent variables are statistically significant. We will revised the model by removing those variables which are not statistically significant (PROX_HAWKER_MARKET, PROX_TOP_PRIMARY_SCH, PROX_KINDERGARTEN, PROX_SUPERMARKET)

Now, we are ready to calibrate the revised model by using the code chunk below.

```{r}
condo.mlr1 <- lm(
  formula = SELLING_PRICE ~ AREA_SQM + AGE + 
    PROX_CBD + PROX_CHILDCARE + PROX_MRT +
    PROX_ELDERLYCARE    + PROX_URA_GROWTH_AREA +
    PROX_PARK   + PROX_PRIMARY_SCH + 
    PROX_SHOPPING_MALL  + PROX_BUS_STOP + 
    NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
  data=condo_resale.sf)
summary(condo.mlr1)
```

The output above reveals that all explanatory variables are statistically significant at 95% confident level.

## 7.5 Prepare Publication Quality Table: gtsummary method

The [**gtsummary**](https://www.danieldsjoberg.com/gtsummary/index.html) package provides an elegant and flexible way to create publication-ready summary tables in R.

In the code chunk below, [`tbl_regression()`](https://www.danieldsjoberg.com/gtsummary/reference/tbl_regression.html) is used to create a well formatted regression report.

```{r}
tbl_regression(condo.mlr1, intercept = TRUE)
```

With gtsummary package, model statistics can be included in the report by either appending them to the report table by using [`add_glance_table()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) or adding as a table source note by using [`add_glance_source_note()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) as shown in the code chunk below.

```{r}
tbl_regression(condo.mlr1, 
               intercept = TRUE) %>% 
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, adj.r.squared, 
                AIC, statistic,
                p.value, sigma))
```

## 7.6 Regression Diagnostics

Regression diagnostics are a set of procedures used to check if a regression model’s assumptions are met and how well the model fits the data. These diagnostics involve checking for issues like non-linear relationships, non-normal errors, non-constant variance, and influential observations to ensure the model’s conclusions are valid and reliable. Common methods include graphical analysis, like residual plots and QQ-plots, and quantitative tests

In this section, we would like to introduce you a fantastic R package specially programmed for performing OLS regression diagnostics. It is called [**olsrr**](https://olsrr.rsquaredacademy.com/). It provides a collection of very useful methods for building better multiple linear regression models:

-   comprehensive regression output
-   residual diagnostics
-   measures of influence
-   heteroskedasticity tests
-   collinearity diagnostics
-   model fit assessment
-   variable contribution assessment
-   variable selection procedures

### 7.6.1 Multicollinearity test

**Multicollinearity** occurs when independent variables are not truly independent, meaning a change in one is associated with a change in another. This makes it hard for the model to isolate each variable’s influence on the outcome.

In the code chunk below, the [`check_collinearity()`](https://easystats.github.io/performance/reference/check_collinearity.html) of **performance** package is used to test if there are sign of multicollinearity.

```{r}
check_collinearity(condo.mlr1)
```

```{r}
#| fig.width: 12
#| fig.height: 10
mlr.vif <- check_collinearity(condo.mlr1)
par(bg="#f6f6f6")
plot(mlr.vif)
```

Since the VIF of the independent variables are less than 10. We can safely conclude that there are no sign of multicollinearity among the independent variables.

### 7.6.2 Test for Non-Linearity

In multiple linear regression, it is important for us to test the assumption that linearity and additivity of the relationship between dependent and independent variables.

In the code chunk below, the [`ols_plot_resid_fit()`](https://olsrr.rsquaredacademy.com/reference/ols_plot_resid_fit.html) of **olsrr** package is used to perform linearity assumption test.

```{r}

check_model(condo.mlr1,
            check = "linearity")
```

The figure above reveals that most of the data poitns are scattered around the 0 line, hence we can safely conclude that the relationships between the dependent variable and independent variables are linear.

### 7.6.3 Test for Normality Assumption

The normality assumption test for multiple linear regression checks **if the model’s residuals (the differences between observed and predicted values) are normally distributed**. This is crucial for accurate hypothesis testing and confidence intervals. To test this, you can use visual methods like histograms and Q-Q plots of the residuals, or conduct statistical tests like Shapiro-Wilk test and Kolmogorov-Smirnov test.

In the code chunk below, `check_normality()` of performance package is used to perform normality assumption test on condo.mlr1 model.

```{r}
check_normality(condo.mlr1)
```

The print above reveals that the p-value of the normality assumption test is less than alpha value of 0.05. Hence we reject the normality assumption at 95% confident level.

::: callout-important
-   `check_normality()` calls *stats::shapiro.test* and checks the standardized residuals (or studentized residuals for mixed models) for normal distribution.
-   Note that this formal test almost always yields significant results for the distribution of residuals and visual inspection (e.g. Q-Q plots) are preferable.
:::

```{r}
res <- data.frame(residuals(condo.mlr1))

ggplot(res, aes(sample = residuals.condo.mlr1.)) +
    stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot of Residuals")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())
```

```{r}
check_model(condo.mlr1,
            check= "normality")
```

The figure reveals that the residual of the multiple linear regression model (i.e. condo.mlr1) is resemble normal distribution.

### 7.6.4 Testing for Spatial Autocorrelation

The hedonic model we try to build are using geographically referenced data. Hence, it is crucial to check for spatial autocorrelation because its presence can produce unreliable and misleading results. Traditional regression models, such as ordinary least squares (OLS), assume that observations are independent of one another. However, spatial data often violates this assumption.

Spatial autocorrelation is the correlation of a variable with itself across different spatial locations. Positive spatial autocorrelation means nearby features tend to be more similar, while negative autocorrelation means they tend to be more dissimilar. This phenomenon is based on the first law of geography: “Everything is related to everything else, but nearby things are more related than distant things”.

**Ignoring spatial autocorrelation in a regression model can lead to serious statistical issues:**

-   Biased and inefficient coefficient estimates: If autocorrelation is present, standard errors of the model coefficients can be wrong, leading to unreliable hypothesis tests (p-values). The model might appear more significant than it is.

-   Misleading significance tests: Standard regression models cannot distinguish between true explanatory power and the influence of spatial patterns, resulting in inaccurate p-values.

-   Model misspecification: Significant spatial autocorrelation in the regression residuals often signals that important explanatory variables are missing from the model. The spatial patterning of the residuals (over- and under-predictions) can provide clues about what these missing variables might be.

-   Inflated Type I error rates: Researchers might incorrectly reject a true null.

To test for spatial autocorrelation, We can run a **Moran’s I test on the model’s residuals**. Significant spatial autocorrelation in the residuals means the model is not capturing the full spatial story.

In order to perform spatial autocorrelation test, we need to export the residual of the hedonic pricing model and save it as a data frame first.

```{r}
mlr.output <- as.data.frame(condo.mlr1$residuals)
```

Next, we will join the newly created data frame with *condo_resale.sf* object.

```{r}
condo_resale.sf <- cbind(condo_resale.sf, 
                        condo.mlr1$residuals) %>%
rename(`MLR_RES` = `condo.mlr1.residuals`)
```

Next, we will use **tmap** package to display the distribution of the residuals on an interactive map.

```{r}
tmap_mode("plot")

```

```{r}
tm_shape(mpsz) +
  tm_polygons(fill_alpha = 0.4) +
tm_shape(condo_resale.sf) +  
  tm_dots(
    fill = "MLR_RES",
    size = 0.3,
    col = "black",
    fill.scale = tm_scale(
      n = 10,
      values = rev(
        brewer.pal(11, "RdBu")),
      style = "quantile",
      midpoint = NA),
    fill.legend = tm_legend(title = "Residuals")
  ) + 
  tm_title("LM Residuals (Quantile Classification)") +
  tm_layout(legend.outside = TRUE) +
  tm_view(set_zoom_limits = c(11,14))
```

```{r}
tmap_mode("plot")
```

The figure above reveal that there is sign of spatial autocorrelation.

To proof that our observation is indeed true, Global Moran’s I test will be performed

First, we will compute the distance-based weight matrix by using [`st_dist_band()`](https://sfdep.josiahparry.com/reference/st_dist_band) function of **sfdep**.

```{r}
condo_resale.sf <- condo_resale.sf %>% 
  mutate(
    nb = st_dist_band(
      st_geometry(geometry), 
                 upper = 1500),
    wt = st_weights(
      nb, style = "W"),
    .before = 1) 
```

Next, [`global_moran_perm()`](https://sfdep.josiahparry.com/reference/global_moran_perm) of **sfdep** package will be used to perform Moran’s I test for residual spatial autocorrelation

```{r}
set.seed(1234)
global_moran_perm(
  condo_resale.sf$MLR_RES, 
  nb = condo_resale.sf$nb, 
  wt = condo_resale.sf$wt,
  alternative = "two.sided", 
  nsim = 499)
```

The Global Moran’s I test for residual spatial autocorrelation shows that it’s p-value is less than the alpha value of 0.05. Hence, we will reject the null hypothesis that the residuals are randomly distributed.

Since the Observed Global Moran I = 0.14389 which is greater than 0, we can infer than the residuals resemble cluster distribution.

# 8 Build Hedonic Pricing Models using GWmodel

In this section, we are going to learn how to modelling hedonic pricing using both the fixed and adaptive bandwidth schemes

## 8.1 Build Fixed Bandwidth GWR Model

### 8.1.1 Compute fixed bandwith

In the code chunk below `bw.gwr()` of GWModel package is used to determine the **optimal fixed bandwidth** to use in the model. Notice that the argument ***adaptive*** is set to **FALSE** indicates that we are interested to compute the fixed bandwidth.

There are two possible approaches can be uused to determine the stopping rule, they are: CV cross-validation approach and AIC corrected (AICc) approach. We define the stopping rule using ***approach*** argeement:

```{r}
bw.fixed <- bw.gwr(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale.sf, 
  approach="CV", 
  kernel="gaussian", 
  adaptive=FALSE, 
  longlat=FALSE)
```

The result shows that the recommended bandwidth is 971.3405 metres since SVY21 is in meter.

### 8.1.2 GWModel method - fixed bandwith

Now we can use the code chunk below to calibrate the gwr model using fixed bandwidth and gaussian kernel:

```{r}
gwr.fixed <- gwr.basic(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale.sf, 
  bw=bw.fixed, 
  kernel = 'gaussian', 
  longlat = FALSE)
```

The output is saved in a list of class “gwrm”. The code below can be used to display the model output.

```{r}
class(gwr.fixed)
```

```{r}
gwr.fixed
```

The report shows that the AICc of the gwr is 42263.61 which is significantly smaller than the globel multiple linear regression model of 42967.1.

## 8.2 Build Adaptive Bandwidth GWR Model

### 8.2.1 Compute the adaptive bandwidth

Similar to the earlier section, we will first use `bw.gwr()` to determine the recommended data point to use.

The code chunk used look very similar to the one used to compute the fixed bandwidth except the `adaptive` argument has changed to **TRUE**.

```{r}
bw.adaptive <- bw.gwr(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + 
    FAMILY_FRIENDLY + FREEHOLD, 
  data=condo_resale.sf, 
  approach="CV", 
  kernel="gaussian", 
  adaptive=TRUE, 
  longlat=FALSE)
```

The result shows that the 30 is the recommended data points to be used.

### 8.2.2 GWModel method - adaptive bandwith

```{r}
gwr.adaptive <- gwr.basic(
  formula = SELLING_PRICE ~ AREA_SQM + AGE + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP +
    NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
  data=condo_resale.sf, 
  bw=bw.adaptive, 
  kernel = 'gaussian', 
  adaptive=TRUE, 
  longlat = FALSE)
```

The code below is used to display the model output.

```{r}
gwr.adaptive
```

The report shows that the AICc the adaptive distance gwr is 41982.22 which is even smaller than the AICc of the fixed distance gwr of 42263.61.

## 8.3 Visualise GWR Output

In addition to regression residuals, the output feature class table includes fields for observed and predicted y values, condition number (cond), Local R2, residuals, and explanatory variable coefficients and standard errors:

-   Condition Number: this diagnostic evaluates local collinearity. In the presence of strong local collinearity, results become unstable. Results associated with condition numbers larger than 30, may be unreliable.

-   Local R2: these values range between 0.0 and 1.0 and indicate how well the local regression model fits observed y values. Very low values indicate the local model is performing poorly. Mapping the Local R2 values to see where GWR predicts well and where it predicts poorly may provide clues about important variables that may be missing from the regression model.

-   Predicted: these are the estimated (or fitted) y values 3. computed by GWR.

-   Residuals: to obtain the residual values, the fitted y values are subtracted from the observed y values. Standardized residuals have a mean of zero and a standard deviation of 1. A cold-to-hot rendered map of standardized residuals can be produce by using these values.

-   Coefficient Standard Error: these values measure the reliability of each coefficient estimate. Confidence in those estimates are higher when standard errors are small in relation to the actual coefficient values. Large standard errors may indicate problems with local collinearity.

They are all stored in a SpatialPointsDataFrame or SpatialPolygonsDataFrame object integrated with fit.points, GWR coefficient estimates, y value, predicted values, coefficient standard errors and t-values in its “data” slot in an object called **SDF** of the output list.

### 8.3.1 Convert SDF into *sf* data.frame

```{r}
condo_resale.sf.adaptive <-
  st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
```

Next, `glimpse()` is used to display the content of *condo_resale.sf.adaptive* sf data frame.

```{r}
glimpse(condo_resale.sf.adaptive)
```

```{r}
summary(gwr.adaptive$SDF$yhat)
```

### 8.3.2 Visualise local R2

```{r}
tm_shape(mpsz)+
  tm_polygons(fill = "white",
    alpha = 0.6,
    border.col = "grey30",
    border.alpha = 0.5) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "Local_R2",
          style = "quantile",
          border.col = "gray60",
          border.lwd = 0.1) +
  tm_layout(bg.color = "#f6f6f6",
            outer.bg.color = "#f6f6f6",
            frame = FALSE,
            legend.outside = TRUE,
            legend.outside.position = "right",
            legend.frame = FALSE)


```

### 8.3.3 Visualise **coefficient estimates**

```{r}
tmap_mode("plot")

AREA_SQM_SE <- tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_SE",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

AREA_SQM_TV <- tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_TV",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_arrange(AREA_SQM_SE, AREA_SQM_TV, 
             asp=1, ncol=2,
             sync = TRUE)

tmap_mode("plot")
```

# 9 Reference

-   Kam, T.S. (2025). [Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method](https://r4gdsa.netlify.app/chap13.html)
