---
title: "Take-home Ex02: Geospatial Analytics for Public Good"
author: Huang Johsuan
date: "2025/10/12"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1 Set the Scene

The locations of businesses and firms are central to urban development, influencing economic growth, job creation, and the availability of resources and services for residents and other businesses. Strategic locations provide competitive advantages, access to talent and infrastructure, and impact brand image and market positioning, while also shaping the city’s overall character and functionality by fostering specialized districts and facilitating efficient supply chains and employee commutes.

Studying the spatio-temporal patterns of businesses and firms is crucial for urban development because it reveals how economic activity and urban forms evolve, enabling better urban planning, promoting sustainability, identifying urban inefficiencies, and enhancing quality of life. By understanding where and when businesses locate and operate, urban planners can design cities more effectively, anticipate growth, ensure resource efficiency, and tailor infrastructure to support economic vitality and community well-being.

# 2 Objective

Utilize **Local Measures of Spatial Autocorrelation (LMSA)** and **Emerging Hot Spot Analysis (EHSA):**

-   To discover how bus mobility patterns differ across neighborhoods, providing deeper insights than global summary measures.
-   To detect emerging patterns in urban mobility, helping planners anticipate commuting demands and design more responsive bus services and policies.

# 3 Tasks

## 3.1 Geospatial Data Science

-   Derive an analytical hexagon data of 375m (this distance is the perpendicular distance between the centre of the hexagon and its edges) to represent the [traffic analysis zone (TAZ)](https://tmg.utoronto.ca/files/Reports/Traffic-Zone-Guidance_March-2021_Final.pdf).

-   With reference to the time intervals provided in the table below, compute the passenger trips generated by origin at the hexagon level.

    | Peak hour period       | Bus tap on time |
    |------------------------|-----------------|
    | Weekday morning peak   | 6am to 9am      |
    | Weekday afternoon peak | 5pm to 8pm      |
    | Weekend/holiday        | 11am to 8pm     |

-   Display the geographical distribution of the passenger trips by using appropriate geovisualisation methods,

-   Describe the spatial patterns revealed by the geovisualisation (not more than 200 words per visual).

## 3.2 LMSA Analysis

-   Compute LMSA statistic of the passengers trips generate by origin at hexagon level.
-   Display the LMSA maps of the passengers trips generate by origin at hexagon level. The maps should only display the significant (i.e. p-value \< 0.05)
-   With reference to the analysis results, draw statistical conclusions (not more than 200 words per visual).

## 3.3 EHSA Analysis

With reference to the passenger trips by origin at the hexagon level for the three time intervals given above:

-   Perform Mann-Kendall Test by using the spatio-temporal local Gi\* values,
-   Prepared EHSA maps of the Gi\* values of the passenger trips by origin at the hexagon level. The maps should only display the significant (i.e. p-value \< 0.05).
-   With reference to the EHSA maps and data visualisation prepared, describe the spatial patterns reveled. (not more than 250 words per cluster).

# 4 The Data

In this exercise, 3 datasets will be used:

-   **Aspatial: Passenger Volume by Origin Destination Bus Stops from LTA DataMall**

    This dataset contains ridership information (datetime, total trips) showing the flow of passengers between bus stops across Singapore's public bus network.

-   **Geospatial: Bus Stop Location from LTA DataMall**

    This dataset provides the geographical coordinates and attributes of all bus stops in Singapore.

-   **Geospatial: Master Plan 2019 Subzone Boundary (No Sea) from** **data.gov.sg**

    This dataset contains polygon boundaries of Singapore's planning subzones as defined in the Urban Redevelopment Authority's (URA) Master Plan 2019.

# 5 Download and Load the Packages

We first use p_load of pacman to download and load the revelent packages:

| Library | Desc |
|----|----|
| [sf](https://r-spatial.github.io/sf/) | Handling geospatial data in R with support for simple features (points, lines, polygons) |
| [tmap](https://r-tmap.github.io/tmap/) | Creating thematic maps and visualization for geospatial data |
| [tidyverse](https://www.tidyverse.org) | A collection of packages for data manipulation and analysis, including readr (reading data), tidyr (tidying data), and dplyr (transforming data) |
| [knitr](https://yihui.org/knitr/) | Generating elegant, flexible and fast dynamic reports that integrate code and output |
| [sfdep](https://sfdep.josiahparry.com) | Creating spatial weights matrix objects from polygon contiguities for spatial dependence analysis |
| [Kendall](https://cran.r-project.org/web/packages/Kendall/refman/Kendall.html#Kendall-package) | Computes the Kendall rank correlation and Mann-Kendall trend test |
| [ggplot2](https://ggplot2.tidyverse.org) | Creating sophisticated data visualizations using the grammar of graphics approach |
| [plotly](https://plotly.com/r/) | Building interactive web-based graphs and visualizations |
| [patchwork](https://patchwork.data-imaginist.com) | Combining multiple ggplot2 plots into a single composition with flexible layout control |

```{r}
pacman::p_load(sf, tmap, tidyverse, knitr,
               sfdep, ggplot2, Kendall, plotly, patchwork,
               ggridges,fs,janitor)
```

# 6 Import & Prepare Data

## 6.1 Import and Transform Geospatial Data

We first import **Master Plan 2019 Subzone Boundary** using `read_rds()` function, and then check its coordinate reference system (CRS) using `st_crs()`:

```{r}
mpsz19 <- read_rds("data/rds/mpsz.rds")
st_crs(mpsz19)
```

The result indicates that the MPSZ19 data is already in the SVY21 coordinate system (EPSG:3414), so no further adjustment is needed.

Next, we import **Bus Stop Location Data** using `st_read()` of sf package and transform the EPSG from 9001 to 3414.

```{r}
BusStop <- st_read(dsn = "data/LTADataMall", 
                 layer = "BusStop") %>%
  st_transform(crs = 3414)

```

Let's check if they are any duplicated Bus stop number:

```{r}
duplicated_ids <- BusStop$BUS_STOP_N[duplicated(BusStop$BUS_STOP_N)]

kable(BusStop%>%
        filter(BUS_STOP_N %in% duplicated_ids)%>%
        arrange(BUS_STOP_N))
```

It seems that the duplicated Bus stop numbers are refering to the same location. To avoid counting same location twice, the first one of the duplicated value will be removed:

```{r}
BusStop <- BusStop %>%
  arrange(BUS_STOP_N, desc(row_number())) %>% # ensures last comes first within each group
  distinct(BUS_STOP_N, .keep_all = TRUE)
```

Let's check if there are any duplicated again.

```{r}
BusStop$BUS_STOP_N[duplicated(BusStop$BUS_STOP_N)]
```

Next, we plot the bus stop locations using **tmap** before proceeding with further analysis:

```{r}
tm_shape(mpsz19)+
  tm_polygons()+
tm_shape(BusStop)+
  tm_dots(size = 0.2)+
  tm_layout(bg.color = "#f6f6f6",frame = FALSE)+
  tm_legend(frame = FALSE)+
  tm_title("Bus Stop Locations",
           position = tm_pos_out("center", "top", "center"))

```

As shown in the map, a few bus stops are located outside Singapore, likely near the Johor Bahru checkpoint in Malaysia. Since our analysis focuses on bus stops within Singapore, these bus stops will be removed from the study area.

The code below uses `st_join()` of **sf** package to clip the study area within SG:

```{r}
BS_SG <- st_join(BusStop,mpsz19,
                         join = st_within, left = FALSE)
```

Let's plot the map and check it again:

```{r}
tm_shape(mpsz19)+
  tm_polygons()+
tm_shape(BS_SG)+
  tm_dots(size = 0.2)+
  tm_layout(bg.color = "#f6f6f6",frame = FALSE)+
  tm_legend(frame = FALSE)+
  tm_title("Bus Stop Locations in Singapore",
           position = tm_pos_out("center", "top", "center"))
```

Now, we can confirm the bus stops located outside of Singapore are all removed.

## 6.2 Import and Prepare Aspatial Data

```{r}
#| eval: false
files <- fs::dir_ls("data/LTADataMall/odbus/") 

# Read all files and clean the column names
odbus <- files %>%
  map_df(~ read_csv(.x, 
                    locale = locale(encoding = "latin1"),
                    col_types = cols(.default = "character")
  ) %>% 
    janitor::clean_names()
  ) 
# Save the output
write.csv(odbus, "data/LTADataMall/odbus/odbus.csv", row.names = FALSE)
```

The code below use `read_csv` to import Jul-Sep 2025 bus trip data:

```{r}
odbus <- read_csv("data/LTADataMall/odbus/odbus.csv")
```

We first view the data and check its data structure:

```{r}
kable(head(odbus))
```

```{r}
str(odbus)
```

As shown in the summary above, this dataset contains 17,306,071 observations and 7 attributes. Although `ORIGIN_PT_CODE` and `DESTINATION_PT_CODE` are already in character format, we need to verify that **each** **bus stop code consists of five digits.**

### 6.2.1 Aspatial Data Preperation

We first check if all bus stop codes are 5-digit using `nchar()` - number of char:

```{r}
all(nchar(odbus$origin_pt_code))
all(nchar(odbus$detination_pt_code))

```

The results indicate that all the numbers are 5 digits long, as they return the value “True”.

The code below renames the attributes for easier reference and combine data later and converts "BUS_STOP_N" into factor for further analysis:

```{r}
#| eval: false
trips <- odbus %>%
  mutate(year_month = ym(year_month),
         Month=month(year_month))%>%
  rename("BUS_STOP_N" = origin_pt_code) %>%
  rename("Day_Type" = day_type) %>%
  rename("Hour" = time_per_hour) %>%
  rename("Trips" = total_trips) %>%
  select(-"year_month",-"pt_type",-"destination_pt_code")

kable(head(trips))
```

```{r}
#| eval: false
write_rds(trips,"data/rds/trip.rds")
```

Lastly, let's check the data summary

```{r}
trips <- read_rds("data/rds/trip.rds")
summary(trips)
```

# 7 EDA of Bus Routes data

## 7.1 Trips Distribution

Let's visualize the data to understand trips distribution through Jul- Sep 2025:

::: panel-tabset
## Plot

```{r}
#| echo: false
hist <- ggplot(trips, aes(x = Trips)) +
  geom_histogram(bins = 30, fill = "#93A4CC", color = "grey30") +
  labs(title = "Histograms of Trips",
       x = "", y = "Frequency") +
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

box <- ggplot(data=trips,aes(x=Trips))+
  geom_boxplot(color="grey30",fill="#93A4CC")+
  labs(
       x = "Trips", y = "") +
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

hist/box +
  plot_annotation(theme = theme(
    plot.background = element_rect(color= '#f6f6f6', linewidth = 0, fill ="#f6f6f6")))
```

## Code

```{r}
#| eval: false
hist <- ggplot(trips, aes(x = Trips)) +
  geom_histogram(bins = 30, fill = "#93A4CC", color = "grey30") +
  labs(title = "Histograms of Hour of Day and Trips",
       x = "", y = "Frequency") +
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

box <- ggplot(data=trips,aes(x=Trips))+
  geom_boxplot(color="grey30",fill="#93A4CC")+
  labs(
       x = "Trips", y = "") +
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

hist/box +
  plot_annotation(theme = theme(
    plot.background = element_rect(color= '#f6f6f6', linewidth = 0, fill ="#f6f6f6")))
```
:::

::: callout-note
## Observation

The distribution of trips is noticeably right-skewed, suggesting that certain bus routes experience a high concentration of activity during specific hours.
:::

## 7.2 Average Trips Distribution by Day of Hour and Day Type

Next, let's check the distribution of trips by hour of day and day type. Before visualization, we need to summarize the average trips by Day_Type and hour first:

```{r}
trips_summary <- trips %>%
  group_by(Day_Type,Month ,Hour) %>%
  summarise(avg_trips = mean(Trips, na.rm = TRUE),.groups = "drop")

```

::: panel-tabset
## Plot

```{r}
#| echo: false
#| fig.width: 12
#| fig.height: 6

trips_summary$Day_Type <- factor(trips_summary$Day_Type, 
                                 levels = c("WEEKENDS/HOLIDAY", "WEEKDAY"))

p <- ggplot(trips_summary, aes(x = Hour, y = Day_Type, fill = Day_Type, 
                                height = avg_trips, group = Day_Type)) +
  geom_ridgeline(scale = 0.1, alpha = 0.7) +
  facet_wrap(~Month, nrow = 3) +  # Show all months at once
  scale_x_continuous(breaks = seq(0, 23, 1)) +
  scale_fill_manual(values = c("#93A4CC", "#8EB859")) +
  labs(
    title = "Average Trips by Month, Hour of Day and Day Type",
    x = "Hour of Day",
    y = ""
  ) +
  geom_vline(xintercept = c(4, 7, 18), color = "grey30", 
             linetype = "dashed", linewidth = 0.3) +
  theme_ridges() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 9),
    axis.title.x = element_text(size = 10,hjust = 0),
    plot.title = element_text(size = 12,hjust = 0.5),
    strip.text = element_text(size = 10),  # Facet title size
    panel.background = element_rect(fill = "#f6f6f6"),
    plot.background = element_rect(fill = "#f6f6f6", color = NA),
    panel.grid = element_blank()
  )

ggplotly(p, tooltip = c("x", "height"),
         )

```

## Code

```{r}
#| eval: false
trips_summary$Day_Type <- factor(trips_summary$Day_Type, 
                                 levels = c("WEEKENDS/HOLIDAY", "WEEKDAY"))

p <- ggplot(trips_summary, aes(x = Hour, y = Day_Type, fill = Day_Type, 
                                height = avg_trips, group = Day_Type)) +
  geom_ridgeline(scale = 0.1, alpha = 0.7) +
  facet_wrap(~Month, nrow = 3) +  # Show all months at once
  scale_x_continuous(breaks = seq(0, 23, 1)) +
  scale_fill_manual(values = c("#93A4CC", "#8EB859")) +
  labs(
    title = "Average Trips by Month, Hour of Day and Day Type",
    x = "Hour of Day",
    y = ""
  ) +
  geom_vline(xintercept = c(4, 7, 18), color = "grey30", 
             linetype = "dashed", linewidth = 0.3) +
  theme_ridges() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 9),
    axis.title.x = element_text(size = 10,hjust = 0),
    plot.title = element_text(size = 12,hjust = 0.5),
    strip.text = element_text(size = 10),  # Facet title size
    panel.background = element_rect(fill = "#f6f6f6"),
    plot.background = element_rect(fill = "#f6f6f6", color = NA),
    panel.grid = element_blank()
  )

ggplotly(p, tooltip = c("x", "height"))
```
:::

::: callout-note
## Observation

On weekdays, three distinct peak periods in bus usage are observed: 04:00 - 05:00 AM, 07:00–08:00 AM, and 18:00–19:00 PM. These peaks are likely driven by commuter traffic, particularly workers traveling to and from their workplaces. (The 4 AM peak is largely due to trips originating from JB into SG, for details, please refer to [8.5 Visualize Total Trips by Hexagon])

During weekends and holidays, the distribution of trips is more flattened throughout the day, with a consistent average number of trips between 07:00 and 19:00.
:::

## 7.3 Number of Bus Stop by Subzone

```{r}
mpsz19$n_bus_stop <-
  lengths(st_intersects(mpsz19, BS_SG))
```

::: panel-tabset
## Plot

```{r}
#| echo: false
tm_shape(mpsz19)+
  tm_polygons(fill = "n_bus_stop",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Number of Bus Stops by Subzone",
           position = tm_pos_out("center", "top", "center"))

  
```

## Code

```{r}
#| eval: false
tm_shape(mpsz19)+
  tm_polygons(fill = "n_bus_stop",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Number of Bus Stops by Subzone",
           position = tm_pos_out("center", "top", "center"))

```
:::

::: callout-note
## Observation

As shown on the map, Murai (87 bus stops), Tampines East (85), and Frankel (80) are the three subzones with the highest number of bus stops.
:::

## 7.4 Trips Volume by Subzone

To examine trip volumes by subzone, we first aggregate the number of trips by bus stop location and then perform a left join with the `mpsz19` sf data for plotting:

```{r}
BS_SG_joined <- BS_SG %>%
   left_join(trips, by = "BUS_STOP_N")

BS_SG_sz <- BS_SG_joined %>%
  st_drop_geometry() %>%  
  select(c("SUBZONE_C","Trips")) %>%
  group_by(SUBZONE_C) %>%
  summarise(total_trips = sum(Trips, na.rm = TRUE))

mpsz19 <- mpsz19 %>% left_join(BS_SG_sz,by = "SUBZONE_C")
```

::: panel-tabset
## Plot

```{r}
#| echo: false

tm_shape(mpsz19)+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Trip Volume by Subzone",
           position = tm_pos_out("center", "top", "center"))

```

## Code

```{r}
#| eval: false

tm_shape(mpsz19)+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Trip Volume by Subzone",
           position = tm_pos_out("center", "top", "center"))
```
:::

Let's print the top 10 subzones with highest trip volume for reference:

```{r}
Top_10_SZ <- mpsz19 %>%
  arrange(desc(total_trips), desc(n_bus_stop)) %>%
  slice(1:10)

kable(Top_10_SZ)
```

::: callout-note
## Observation

Tampines East appears on the list again, indicating that it has both a high number of bus stops and significant traffic volume. Observing the other subzones on the list, most contain major bus interchanges, suggesting that these areas serve as key transit hubs.
:::

# 8 Derive and Visualize Analytical Hexagons

One of the task is to derive an analytical hexagon data of 375m (this distance is the perpendicular distance between the centre of the hexagon and its edges) to represent the [traffic analysis zone (TAZ)](https://tmg.utoronto.ca/files/Reports/Traffic-Zone-Guidance_March-2021_Final.pdf).

According to the documentation, the cell size of hexagonal cells refers to the distance between opposite edges. Therefore, the cell size in this case is 750 m.

The code below uses st_make_grid() to dervie the hexagon for TAZ:

```{r}
hexagon <- st_make_grid(mpsz19,
                        cellsize = 750, 
                        what = "polygon",
                        square = FALSE) %>% 
  st_sf()

hexagon
```

```{r}
kable(head(hexagon))
```

Notice that the hexagon contains only one column with polygon geometry, lacking an unique ID. We need to map the hexagons with bus stops:

```{r}
hexagon$n_collisions <-
  lengths(st_intersects(hexagon, BS_SG))

kable(head(hexagon))
```

## 8.1 Filter Hexagons for Study Area

Next, we filter `n_collisions` greater than 0 for the study area:

```{r}
hexagon <- filter(hexagon,n_collisions>0)
count(hexagon)
```

There are total 834 hexagon as our study area.

## 8.2 Assign IDs to Each Hexagon

We need an unique identifier for each hexagon for reference. The code below defines and assigns new ID formatting in Hxxx for each hexagon:

```{r}
hexagon <- hexagon %>%
  select(,-n_collisions) # "-" means drop the n_collisions column

hexagon$HEX_ID <- sprintf(
  "H%04d", seq_len(nrow(hexagon))) %>%
  as.factor()
```

Next, we plot our study area using tmap:

```{r}
tm_shape(mpsz19)+
  tm_polygons()+
  tm_shape(hexagon)+
  tm_fill(col= "black",fill="#cc78aa")+
  tm_layout(bg.color = "#f6f6f6",frame = FALSE)+
  tm_legend(frame = FALSE)+
  tm_title("Traffic Analysis Zone",
           position = tm_pos_out("center", "top", "center"))
```

## 8.3 Combine Hexagon ID with Trip Data

We first map Bus Stop code with Hexagon ID:

```{r}
bus_hex <- st_intersection(
  BS_SG, hexagon) %>%
  st_drop_geometry() %>%
  select(c(BUS_STOP_N, HEX_ID))%>%
  distinct(BUS_STOP_N, .keep_all = TRUE)


kable(head(bus_hex))
```

Next, we use `inner_join()` to include the trips data within hexagon study areas:

```{r}
trips <- trips %>% 
  inner_join(bus_hex, by = "BUS_STOP_N")
```

## 8.4 Compute Total Trips by Peak Hour Period and Hexagon

Following the task instructions, the `peak` variable was created using the `mutate()` and `case_when()` functions. However, based on the earlier EDA, where a peak was also observed between 4 AM and 5 AM, this time period has been included as part of the peak definition.

```{r}
trips <- trips %>%
  mutate(
    Peak = case_when(
      Day_Type == "WEEKDAY" & Hour >= 4 & Hour < 5  ~ "Weekday 4 AM Peak",
      Day_Type == "WEEKDAY" & Hour >= 6 & Hour < 9  ~ "Weekday AM Peak",
      Day_Type == "WEEKDAY" & Hour >= 17 & Hour < 20 ~ "Weekday PM Peak",
      Day_Type == "WEEKENDS/HOLIDAY" & Hour >= 11 & Hour < 20 ~ "Weekend/Holiday Peak",
      TRUE ~ "Off Peak"
    )
  )
```

Next, we sum up the total trips by HEX_ID and Peak hour period:

```{r}
trips_summary_hex <- trips %>%
  group_by(HEX_ID, Peak) %>%
  summarise(total_trips = sum(Trips, na.rm = TRUE), .groups = "drop") %>%  # drop grouping here
  left_join(hexagon, by = "HEX_ID")%>%
  st_sf()
```

## 8.5 Visualize Total Trips by Hexagon

```{r}
ggplot(data = trips_summary_hex) +
  geom_histogram(aes(x = total_trips), bins = 30, fill = "#93A4CC", color = "grey30") +
  labs(
    title = "Distribution of Total Trips by Hexagon",
    x = "Total Trips",
    y = "Count"
  ) +
   theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())

```

Since the total trips during peak hours are still fairly left-skewed, we use "jenks" natural break method to highlight areas with disproportionately high values:

::: panel-tabset
## Plot

```{r}
#| echo: false
#| fig-height: 12
#| fig-width: 12

WD_4AM <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekday 4 AM Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekday AM Peak: 4-5",
           position = tm_pos_out("center", "top", "center"))

WK <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekend/Holiday Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekend/Holiday Peak: 11-20",
           position = tm_pos_out("center", "top", "center"))

WD_AM <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekday AM Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekday AM Peak: 6-9",
           position = tm_pos_out("center", "top", "center"))

WD_PM <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekday PM Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekday PM Peak: 17-20",
           position = tm_pos_out("center", "top", "center"))



tmap_arrange(WD_4AM,WD_AM, WD_PM,WK,ncol = 2,nrow=2)
```

## Code

```{r}
#| eval: false

WD_4AM <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekday 4 AM Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekday AM Peak: 4-5",
           position = tm_pos_out("center", "top", "center"))

WK <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekend/Holiday Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekend/Holiday Peak: 11-20",
           position = tm_pos_out("center", "top", "center"))

WD_AM <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekday AM Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekday AM Peak: 6-9",
           position = tm_pos_out("center", "top", "center"))

WD_PM <- tm_shape(mpsz19)+
  tm_polygons(fill_alpha = 0.5,lwd =0.5,col_alpha=0.5)+
tm_shape(trips_summary_hex %>% filter (Peak=="Weekday PM Peak"))+
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(style = "jenks",
                                              n=5))+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_legend(frame = FALSE)+
  tm_title("Weekday PM Peak: 17-20",
           position = tm_pos_out("center", "top", "center"))



tmap_arrange(WD_4AM,WD_AM, WD_PM,WK,ncol = 2,nrow=2)
```
:::

Next, let's list out the top 5 Hexagon during each peak period:

```{r}
trips_summary_hex <- st_join(trips_summary_hex,mpsz19,
                         join = st_within, left = FALSE)
```

```{r}
kable(
  trips_summary_hex %>%
    filter(Peak %in% c("Weekday AM Peak", "Weekday PM Peak", "Weekend/Holiday Peak")) %>%
    group_by(Peak) %>%
    slice_max(order_by = total_trips.x, n = 5, with_ties = FALSE) %>%
    ungroup() %>%
    select(HEX_ID, SUBZONE_N, Peak, total_trips.x) %>%
    arrange(Peak, desc(total_trips.x))
)
  
```

::: callout-note
## Observation

According to [The Stait Time](https://www.straitstimes.com/singapore/singapores-cross-border-bus-services-to-start-leaving-jb-checkpoint-earlier-at-4-50am-on-weekdays), starting from Sep 15, the first buses of four Singapore-operated services departing from the Johor Bahru checkpoint to Woodlands on weekdays will begin operations earlier at 4:50 AM to help alleviate early morning traffic congestion at the Causeway.

Since the plot above only includes bus trips originating from SG, it only partially reflects the early morning peak observed in the previous visualizations. Most buses operating around 4 AM originate from the JB checkpoint rather than from within Singapore.

During both regular weekday and weekend peak hours, the busiest areas based on total trips are hexagons H0110, H0768, H0350, and H0400. These are located in the westernmost (Jurong West Central), easternmost (Tampines East), and northernmost regions (Sembawang Central and Woodlands East). Notably, H0667 (Geylang East) ranks as the third busiest area during weekend and holiday peak hours. In these areas, commuters are more likely to rely on buses due to limited or absent train service.
:::

# 9 LMSA Analysis - HCSA

To support the LMSA analysis, Hot Spot/Cold Spot analysis (HCSA) based on the Getis-Ord Gi\* statistic will be performed.

```{r}
trips_hex_hcsa <- trips %>%
  group_by(HEX_ID,Day_Type)%>%
  summarise(total_trips = sum(Trips,na.rm = TRUE),.groups = "drop")

trips_hex_hcsa <- hexagon%>%
  inner_join(trips_hex_hcsa, by = "HEX_ID")
```

## 9.1 Dervice Adaptive Distance Weight Matrix

Since some of the hexagons on the rightmost of Singapore are isolated, and no neighbors directly connected on the edges. The **adaptive distance weight matrix** - KNN method ensure every hexagon has neighbors.

The code chunk below will be used to derive a KNN weights:

nb = include_self(st_knn(geometry,k=6)),

wt = st_weights(nb,style = "W"),

.before = 1)

```{r}
wm_knn_wk <- trips_hex_hcsa %>%
  filter(Day_Type=="WEEKENDS/HOLIDAY") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1)

wm_knn_wd <- trips_hex_hcsa %>%
  filter(Day_Type=="WEEKDAY") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1)
```

## 9.2 Compute Gi\*

The code below uses `mutate()` to add a new column called `gistar`, which stores the results of `local_gstar_perm()`. This function computes the local Gi\* statistic, a spatial measure used to identify local clusters of high values (hot spots) and low values (cold spots).

The hypothesis of `local_gstar_perm()`:

-   **H0:** There is no spatial clustering of the variable of interest.
-   **H1:** There is spatial clustering (hot spot or cold spot) at a location.
-   **Significance**: 5% (since this is a two-sided test, the critical region is split between the two tails of the distribution, each side is 2.5%)

```{r}
set.seed(1234)

trips_hcsa_wk <- wm_knn_wk %>%
  mutate(
    gistar = local_gstar_perm(
      total_trips, nb, wt, nsim = 99,
      alternative = "two.sided"),
    .before = 1) %>%
  unnest(gistar)

trips_hcsa_wd <- wm_knn_wd %>%
  mutate(
    gistar = local_gstar_perm(
      total_trips, nb, wt, nsim = 99,
      alternative = "two.sided"),
    .before = 1) %>%
  unnest(gistar)
```

## 9.2 **Visualize local HCSA**

Here we visualize cold/hot spots using a two-sided test, filtering locations with folded simulated p-values \< 0.025, where:

-   **Gi\* \> 0 & `p_sim_folded` \<= 0.025 → Hot spot**
-   **Gi\* \< 0 & `p_sim_folded`\<= 0.025 → Cold spot**

```{r}
#| fig.width: 12
#| fig.height: 6

wk <- tm_shape(mpsz19)+
  tm_fill(fill_alpha = 0.5)+
  tm_shape(hexagon) + 
  tm_borders(col = "grey30", col_alpha=0.8)+
tm_shape(trips_hcsa_wk %>% filter(p_folded_sim <= 0.025))+
  tm_fill("cluster",
          fill.scale = tm_scale_categorical(
               values= c("#93A4CC","#cc78aa")
              ))+
  tm_title("Hot Spot and Cold Spot on Weenkends",
           position = tm_pos_out("center", "top", "center"))+
  tm_layout(
    bg.color = "#f6f6f6",
    frame = FALSE)

wd <- 
  tm_shape(mpsz19)+
  tm_fill(fill_alpha = 0.5)+
  tm_shape(hexagon) + 
  tm_borders(col = "grey30", col_alpha=0.8)+
tm_shape(trips_hcsa_wd %>% filter(p_folded_sim <= 0.025))+
  tm_fill("cluster",
          fill.scale = tm_scale_categorical(
               values= c("#93A4CC","#cc78aa"))
          )+
  tm_title("Hot Spot and Cold Spot on Weekdays",
           position = tm_pos_out("center", "top", "center"))+
  tm_layout(
    bg.color = "#f6f6f6",
    frame = FALSE)

tmap_arrange(wd,wk,ncol=2)
```

::: callout-note
## Observation
:::

# 10 EHSA Analysis

## 10.1 By Month

### 10.1.1 Data Preparation for EHSA Analysis

```{r}
trips_morning_m <- trips %>%
  filter(Peak == "Weekday AM Peak") %>%
  group_by(HEX_ID, Month) %>%
  summarise(Total_trips = sum(Trips, na.rm = TRUE), .groups = "drop")

trips_afternoon_m <- trips %>%
  filter(Peak == "Weekday PM Peak") %>%
  group_by(HEX_ID, Month) %>%
  summarise(Total_trips = sum(Trips, na.rm = TRUE), .groups = "drop")

trips_weekend_m <- trips %>%
  filter(Peak == "Weekend/Holiday Peak") %>%
  group_by(HEX_ID, Month) %>%
  summarise(Total_trips = sum(Trips, na.rm = TRUE), .groups = "drop")

# full grid for space time
full_grid <- expand.grid(
  HEX_ID = hexagon$HEX_ID,
  Month = 7:9
)

# combine 
trips_morning_m <- full_grid %>%
  left_join(trips_morning_m, by = c("HEX_ID", "Month")) %>%
  mutate(Total_trips = ifelse(is.na(Total_trips), 0, Total_trips))

trips_afternoon_m <- full_grid %>%
  left_join(trips_afternoon_m, by = c("HEX_ID", "Month")) %>%
  mutate(Total_trips = ifelse(is.na(Total_trips), 0, Total_trips))

trips_weekend_m <- full_grid %>%
  left_join(trips_weekend_m, by = c("HEX_ID", "Month")) %>%
  mutate(Total_trips = ifelse(is.na(Total_trips), 0, Total_trips))
```

### 10.1.2 Create a Time Series Cube

```{r}
trips_morning_st_m <- spacetime(trips_morning_m,hexagon,
                      .loc_col = "HEX_ID",
                      .time_col = "Month")
is_spacetime_cube(trips_morning_st_m)
```

```{r}
trips_afternoon_st_m <- spacetime(trips_afternoon_m,hexagon,
                      .loc_col = "HEX_ID",
                      .time_col = "Month")
is_spacetime_cube(trips_afternoon_st_m)
```

```{r}
trips_weekend_st_m <- spacetime(trips_weekend_m,hexagon,
                      .loc_col = "HEX_ID",
                      .time_col = "Month")
is_spacetime_cube(trips_weekend_st_m)
```

### 10.1.3 Derive Adaptive Distance Weight Matrix

```{r}
trips_morning_nb_m <- trips_morning_st_m %>%
  activate("geometry") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")

trips_afternoon_nb_m <- trips_afternoon_st_m %>%
  activate("geometry") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")

trips_weekend_nb_m <- trips_weekend_st_m %>%
  activate("geometry") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")
```

### 10.1.4 Calculate Gi\*

```{r}
nb <- include_self(st_knn(st_geometry(hexagon), k = 6))
wt <- st_weights(nb, style = "W")

trips_morning_spatial <- hexagon %>%
  left_join(trips_morning_m, by = "HEX_ID")

trips_afternoon_spatial <- hexagon %>%
  left_join(trips_afternoon_m, by = "HEX_ID")

trips_weekend_spatial <- hexagon %>%
  left_join(trips_weekend_m, by = "HEX_ID")
```

```{r}
gi_stars_morning <- trips_afternoon_spatial %>%
  group_by(Month) %>%
  mutate(
    gi_star = local_gstar_perm(
      Total_trips,
      nb = nb,
      wt = wt,
      nsim = 99,
      alternative = "two.sided"
    )
  ) %>%
  unnest(gi_star) %>%
  ungroup()

gi_stars_afternoon <- trips_afternoon_spatial %>%
  group_by(Month) %>%
  mutate(
    gi_star = local_gstar_perm(
      Total_trips,
      nb = nb,
      wt = wt,
      nsim = 99,
      alternative = "two.sided"
    )
  ) %>%
  unnest(gi_star) %>%
  ungroup()

gi_stars_weekend <- trips_weekend_spatial %>%
  group_by(Month) %>%
  mutate(
    gi_star = local_gstar_perm(
      Total_trips,
      nb = nb,
      wt = wt,
      nsim = 99,
      alternative = "two.sided"
    )
  ) %>%
  unnest(gi_star) %>%
  ungroup()
```

### 10.1.5 **Mann-Kendall Test on Gi\***

A **monotonic series** or function is one that only increases (or decreases) and never changes direction. So long as the function either stays flat or continues to increase, it is monotonic.

-   H0: No monotonic trend
-   H1: Monotonic trend is present

```{r}
mk_results_m <- gi_stars_morning %>%
  group_by(HEX_ID) %>%
  summarise(
    mk_tau = MannKendall(gi_star)$tau,
    mk_pvalue = MannKendall(gi_star)$sl,
    trend_direction = case_when(
      mk_pvalue > 0.05 ~ "No significant trend",
      mk_tau > 0 ~ "Increasing trend",
      mk_tau < 0 ~ "Decreasing trend"
    ),
    .groups = "drop"
  )

mk_results_a <- gi_stars_afternoon %>%
  group_by(HEX_ID) %>%
  summarise(
    mk_tau = MannKendall(gi_star)$tau,
    mk_pvalue = MannKendall(gi_star)$sl,
    trend_direction = case_when(
      mk_pvalue > 0.05 ~ "No significant trend",
      mk_tau > 0 ~ "Increasing trend",
      mk_tau < 0 ~ "Decreasing trend"
    ),
    .groups = "drop"
  )

mk_results_wk <- gi_stars_weekend%>%
  group_by(HEX_ID) %>%
  summarise(
    mk_tau = MannKendall(gi_star)$tau,
    mk_pvalue = MannKendall(gi_star)$sl,
    trend_direction = case_when(
      mk_pvalue > 0.05 ~ "No significant trend",
      mk_tau > 0 ~ "Increasing trend",
      mk_tau < 0 ~ "Decreasing trend"
    ),
    .groups = "drop"
  )
```

```{r}
mk_results_m %>%
  group_by(trend_direction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

mk_results_a %>%
  group_by(trend_direction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

mk_results_wk %>%
  group_by(trend_direction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

### 10.1.6 Visualize MK test results

Since

```{r}
tm_shape(mpsz19)+
  tm_fill(fill_alpha = 0.5)+
  tm_shape(hexagon) + 
  tm_borders(col = "grey30", col_alpha=0.8)+
tm_shape(mk_results_wk)+
  tm_fill("trend_direction",
          fill.scale = tm_scale_categorical(
               values= c("#93A4CC","#cc78aa","#f6f6f6")
              ))+
  tm_title("Hot Spot and Cold Spot on Weenkends",
           position = tm_pos_out("center", "top", "center"))+
  tm_layout(
    bg.color = "#f6f6f6",
    frame = FALSE)
```

```{r}
# Suppose mk_results_wd has HEX_ID, Year, gi_star
hex_ids <- unique(gi_stars_weekend$HEX_ID)

# Create traces for each HEX_ID
fig <- plot_ly()
for (id in hex_ids) {
  df <- gi_stars_morning %>% filter(HEX_ID == id)
  fig <- fig %>% add_trace(
    data = df,
    x = ~Month,
    y = ~gi_star,
    type = 'scatter',
    mode = 'lines',
    name = id,
    visible = ifelse(id == hex_ids[1], TRUE, FALSE)
  )
}

# Dropdown menu
buttons <- lapply(seq_along(hex_ids), function(i) {
  visible <- rep(FALSE, length(hex_ids))
  visible[i] <- TRUE
  list(
    method = "update",
    args = list(list(visible = visible),
                list(title = paste("HEX_ID:", hex_ids[i]))),
    label = hex_ids[i]
  )
})

fig <- fig %>% layout(
  updatemenus = list(list(
    active = 0,
    buttons = buttons,
    x = 0.1,
    y = 1.15
  )),
  title = paste("HEX_ID:", hex_ids[1]),
  xaxis = list(title = "Month"),
  yaxis = list(title = "Gi*")
)

fig
```

### 10.1.7 HCSA

```{r}
ehsa_morning <- emerging_hotspot_analysis(
  x = trips_morning_nb, 
  include_gi = TRUE,
  .var = "Total_trips", 
  k = 1,      # time lag = 1
  nsim = 199, # 200 simulation
  nb_col = "nb",    
  wt_col = "wt",
  threshold = 0.05
)

ehsa_afternoon <- emerging_hotspot_analysis(
  x = trips_afternoon_nb, 
  include_gi = TRUE,
  .var = "Total_trips", 
  k = 1,      # time lag = 1
  nsim = 199, # 200 simulation
  nb_col = "nb",    
  wt_col = "wt",
  threshold = 0.05
)

ehsa_weekend <- emerging_hotspot_analysis(
  x = trips_afternoon_nb, 
  include_gi = TRUE,
  .var = "Total_trips", 
  k = 1,      # time lag = 1
  nsim = 199, # 200 simulation
  nb_col = "nb",    
  wt_col = "wt",
  threshold = 0.05
)
```

```{r}
ehsa_morning <- hexagon %>%
  left_join(ehsa_morning, by = join_by(HEX_ID == location))
```

```{r}
ehsa_sig <- ehsa_morning  %>%
  filter(p_value < 0.05)


tm_shape(ehsa_morning) +
  tm_polygons() +
  tm_borders(fill_alpha = 0.5) +
tm_shape(ehsa_sig) +
  tm_fill("classification") + 
  tm_borders(fill_alpha = 0.4)+
  tm_title("Emerging Hotspot Analysis",
           position = tm_pos_out("center", "top", "center"))+
  tm_layout(
    bg.color = "#f6f6f6",
    frame = FALSE)
```

## 10.2 By Hour

### 10.1.1 Data Preparation for EHSA Analysis

```{r}
trips_morning <- trips %>%
  filter(Peak == "Weekday AM Peak") %>%
  group_by(HEX_ID, Hour) %>%
  summarise(Total_trips = sum(Trips, na.rm = TRUE), .groups = "drop")

trips_afternoon <- trips %>%
  filter(Peak == "Weekday PM Peak") %>%
  group_by(HEX_ID, Hour) %>%
  summarise(Total_trips = sum(Trips, na.rm = TRUE), .groups = "drop")

trips_weekend <- trips %>%
  filter(Peak == "Weekend/Holiday Peak") %>%
  group_by(HEX_ID, Hour) %>%
  summarise(Total_trips = sum(Trips, na.rm = TRUE), .groups = "drop")

# full grid for space time
full_grid_morning <- expand.grid(
  HEX_ID = hexagon$HEX_ID,
  Hour = 6:9
)

full_grid_afternoon <- expand.grid(
  HEX_ID = hexagon$HEX_ID,
  Hour = 17:20
)

full_grid_weekend <- expand.grid(
  HEX_ID = hexagon$HEX_ID,
  Hour = 11:20
)

# combine 
trips_morning <- full_grid_morning %>%
  left_join(trips_morning, by = c("HEX_ID", "Hour")) %>%
  mutate(Total_trips = ifelse(is.na(Total_trips), 0, Total_trips))

trips_afternoon <- full_grid_afternoon %>%
  left_join(trips_afternoon, by = c("HEX_ID", "Hour")) %>%
  mutate(Total_trips = ifelse(is.na(Total_trips), 0, Total_trips))

trips_weekend <- full_grid_weekend %>%
  left_join(trips_weekend, by = c("HEX_ID", "Hour")) %>%
  mutate(Total_trips = ifelse(is.na(Total_trips), 0, Total_trips))
```

### 10.1.2 Create a Time Series Cube

```{r}
trips_morning_st <- spacetime(trips_morning,hexagon,
                      .loc_col = "HEX_ID",
                      .time_col = "Hour")
is_spacetime_cube(trips_morning_st)
```

```{r}
trips_afternoon_st <- spacetime(trips_afternoon,hexagon,
                      .loc_col = "HEX_ID",
                      .time_col = "Hour")
is_spacetime_cube(trips_afternoon_st)
```

```{r}
trips_weekend_st <- spacetime(trips_weekend,hexagon,
                      .loc_col = "HEX_ID",
                      .time_col = "Hour")
is_spacetime_cube(trips_weekend_st)
```

### 10.1.3 Derive Adaptive Distance Weight Matrix

```{r}
trips_morning_nb_2 <- trips_morning_st %>%
  activate("geometry") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")

trips_afternoon_nb_2 <- trips_afternoon_st %>%
  activate("geometry") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")

trips_weekend_nb_2 <- trips_weekend_st %>%
  activate("geometry") %>%
  mutate(
    nb = include_self(st_knn(geometry,k=6)),
    wt = st_weights(nb,style = "W"),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")
```

### 10.1.4 Calculate Gi\*

```{r}
nb <- include_self(st_knn(st_geometry(hexagon), k = 6))
wt <- st_weights(nb, style = "W")

trips_morning_spatial <- hexagon %>%
  left_join(trips_morning, by = "HEX_ID")

trips_afternoon_spatial <- hexagon %>%
  left_join(trips_afternoon, by = "HEX_ID")

trips_weekend_spatial <- hexagon %>%
  left_join(trips_weekend, by = "HEX_ID")
```

```{r}
gi_stars_morning <- trips_afternoon_spatial %>%
  group_by(Hour) %>%
  mutate(
    gi_star = local_gstar_perm(
      Total_trips,
      nb = nb,
      wt = wt,
      nsim = 99,
      alternative = "two.sided"
    )
  ) %>%
  unnest(gi_star) %>%
  ungroup()

gi_stars_afternoon <- trips_afternoon_spatial %>%
  group_by(Hour) %>%
  mutate(
    gi_star = local_gstar_perm(
      Total_trips,
      nb = nb,
      wt = wt,
      nsim = 99,
      alternative = "two.sided"
    )
  ) %>%
  unnest(gi_star) %>%
  ungroup()

gi_stars_weekend <- trips_weekend_spatial %>%
  group_by(Hour) %>%
  mutate(
    gi_star = local_gstar_perm(
      Total_trips,
      nb = nb,
      wt = wt,
      nsim = 99,
      alternative = "two.sided"
    )
  ) %>%
  unnest(gi_star) %>%
  ungroup()
```

### 10.1.5 **Mann-Kendall Test on Gi\***

A **monotonic series** or function is one that only increases (or decreases) and never changes direction. So long as the function either stays flat or continues to increase, it is monotonic.

-   H0: No monotonic trend
-   H1: Monotonic trend is present

```{r}
mk_results_m <- gi_stars_morning %>%
  group_by(HEX_ID) %>%
  summarise(
    mk_tau = MannKendall(gi_star)$tau,
    mk_pvalue = MannKendall(gi_star)$sl,
    trend_direction = case_when(
      mk_pvalue > 0.05 ~ "No significant trend",
      mk_tau > 0 ~ "Increasing trend",
      mk_tau < 0 ~ "Decreasing trend"
    ),
    .groups = "drop"
  )

mk_results_a <- gi_stars_afternoon %>%
  group_by(HEX_ID) %>%
  summarise(
    mk_tau = MannKendall(gi_star)$tau,
    mk_pvalue = MannKendall(gi_star)$sl,
    trend_direction = case_when(
      mk_pvalue > 0.05 ~ "No significant trend",
      mk_tau > 0 ~ "Increasing trend",
      mk_tau < 0 ~ "Decreasing trend"
    ),
    .groups = "drop"
  )

mk_results_wk <- gi_stars_weekend%>%
  group_by(HEX_ID) %>%
  summarise(
    mk_tau = MannKendall(gi_star)$tau,
    mk_pvalue = MannKendall(gi_star)$sl,
    trend_direction = case_when(
      mk_pvalue > 0.05 ~ "No significant trend",
      mk_tau > 0 ~ "Increasing trend",
      mk_tau < 0 ~ "Decreasing trend"
    ),
    .groups = "drop"
  )
```

```{r}
mk_results_m %>%
  group_by(trend_direction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

mk_results_a %>%
  group_by(trend_direction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

mk_results_wk %>%
  group_by(trend_direction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

### 10.1.6 Visualize MK test results

Since

```{r}
tm_shape(mpsz19)+
  tm_fill(fill_alpha = 0.5)+
  tm_shape(hexagon) + 
  tm_borders(col = "grey30", col_alpha=0.8)+
tm_shape(mk_results_wk)+
  tm_fill("trend_direction",
          fill.scale = tm_scale_categorical(
               values= c("#93A4CC","#cc78aa","#f6f6f6")
              ))+
  tm_title("Hot Spot and Cold Spot on Weenkends",
           position = tm_pos_out("center", "top", "center"))+
  tm_layout(
    bg.color = "#f6f6f6",
    frame = FALSE)
```

```{r}
# Suppose mk_results_wd has HEX_ID, Year, gi_star
hex_ids <- unique(gi_stars_weekend$HEX_ID)

# Create traces for each HEX_ID
fig <- plot_ly()
for (id in hex_ids) {
  df <- gi_stars_morning %>% filter(HEX_ID == id)
  fig <- fig %>% add_trace(
    data = df,
    x = ~Hour,
    y = ~gi_star,
    type = 'scatter',
    mode = 'lines',
    name = id,
    visible = ifelse(id == hex_ids[1], TRUE, FALSE)
  )
}

# Dropdown menu
buttons <- lapply(seq_along(hex_ids), function(i) {
  visible <- rep(FALSE, length(hex_ids))
  visible[i] <- TRUE
  list(
    method = "update",
    args = list(list(visible = visible),
                list(title = paste("HEX_ID:", hex_ids[i]))),
    label = hex_ids[i]
  )
})

fig <- fig %>% layout(
  updatemenus = list(list(
    active = 0,
    buttons = buttons,
    x = 0.1,
    y = 1.15
  )),
  title = paste("HEX_ID:", hex_ids[1]),
  xaxis = list(title = "Hour"),
  yaxis = list(title = "Gi*")
)

fig
```

### 10.1.7 HCSA

```{r}

ehsa_weekend <- emerging_hotspot_analysis(
  x = trips_weekend_nb_2, 
  include_gi = TRUE,
  .var = "Total_trips", 
  k = 1,      # time lag = 1
  nsim = 199, # 200 simulation
  nb_col = "nb",    
  wt_col = "wt",
  threshold = 0.05
)
```

```{r}
ggplot(data = ehsa_morning,
       aes(x = classification)) +
  geom_bar(fill = "#93A4CC", color = "grey30")+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "#f6f6f6"),
        plot.background = element_rect(fill = "#f6f6f6",color = NA),
        panel.grid = element_blank())
```

```{r}

```
