---
title: "Hands-on 3: Spatio-Temporal Point Patterns Analysis"
author: Johsuan Huang
date: 2025-09-09
date-modified: "last-modified"
description: ""
format: html
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

A spatio-temporal point process (also called space-time or spatial-temporal point process) is a random collection of points, where each point represents the time and location of an event. Examples of events include incidence of disease, sightings or births of a species, or the occurrences of fires, earthquakes, lightning strikes, tsunamis, or volcanic eruptions.

The analysis of spatio-temporal point patterns is becoming increasingly necessary, given the rapid emergence of geographically and temporally indexed data in a wide range of fields. Several spatio-temporal point patterns analysis methods have been introduced and implemented in R in the last ten years. This chapter shows how various R packages can be combined to run a set of spatio-temporal point pattern analyses in a guided and intuitive way. A real world forest fire events in Kepulauan Bangka Belitung, Indonesia from 1st January 2023 to 31st December 2023 is used to illustrate the methods, procedures and interpretations.

## 1.1 The Research Questions

-   Are the locations of forest fire in Kepulauan Bangka Belitung spatial and spatio-temporally independent?
-   If the answer is NO, where and when the observed forest fire locations tend to cluster?

# 2 The Data

To provide answers to the questions above, two data sets will be used. They are:

-   ***forestfires***, a csv file provides locations of forest fire detected from the Moderate Resolution Imaging Spectroradiometer (MODIS) sensor data. The data are downloaded from [Fire Information for Resource Management System](https://firms.modaps.eosdis.nasa.gov/download/). For the purpose of this exercise, only forest fires within Kepulauan Bangka Belitung will be used.

-   ***Kepulauan_Bangka_Belitung***, an ESRI shapefile showing the sub-district (i.e. [kelurahan](https://en.wikipedia.org/wiki/Villages_of_Indonesia)) boundary of Kepulauan Bangka Belitung. The data set was downloaded from [Indonesia Geospatial](https://www.indonesia-geospasial.com/2023/05/download-shapefile-batas-administrasi.html) portal. The original data covers the whole Indonesia. For the purpose of this exercise, only sub-districts within Kepulauan Bangka Belitung are extracted.

# 3 Import Libraries

| Library | Desc |
|----------------|----------------------------------------------|
| [**sf**](https://r-spatial.github.io/sf/) | handling geospatial data |
| [**tmap**](https://cran.r-project.org/web/packages/tmap/index.html) | visualization for geospatial data |
| **tidyverse** | handling data including packages **readr**, **tidyr and dplyr** |
| [**sparr**](https://tilmandavies.github.io/sparr/index.html) | provides functions to estimate fixed and adaptive kernel-smoothed spatial relative risk surfaces via the density-ratio method and perform subsequent inference.  |
| [**spatstat**](https://spatstat.org/) | for point pattern analysis |

```{r}
pacman::p_load(sf, spatstat,tmap, tidyverse,knitr,sparr)
```

# 4 Import Data

## 4.1 Import Study Area

We use `st_union()` to remove all the sub-borders and merge them into one big polygon:

```{r}
# The raw file is WGS84
kbb_sf <- st_read(dsn="Data/rawdata",
               layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>% 
  st_zm(drop = TRUE, what="ZM") %>%
  st_transform(crs = 32748)
  
```

```{r}
tm_shape(kbb_sf)+
  tm_polygons()+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")+
  tm_title("Kepulauan Bangka Belitung")
      
```

::: callout-note
## UTM 32748

-   **Type**: Projected coordinate systems
-   **Coverage:** Indonesia and parts of Southeast Asia
-   **Zone:** UTM Zone 48 South (102°E to 108°E, south of equator)
-   **Use:** Regional projection for Indonesia
-   **Units:** Meters
:::

### 4.1.1 Converting to OWIN object

Next, `as.owin()` is used to convert *kbb* into an **owin** object required by **spatstat**

```{r}
kbb_owin = as.owin(kbb_sf)
kbb_owin
```

`class()` is used to confirm if the output is indeed an owin object:

```{r}
class(kbb_owin)
```

## 4.2 Import & Prepare Forest Fire Data

The dataset documents forest fires detected by satellites between 2023-01-10 and 2023-12-18, comprising of 741 observations.

```{r}
fire_sf <- read_csv("Data/forestfires.csv") %>%
  st_as_sf(coords=c("longitude", "latitude"), crs = 4326) %>% #raw GPS coordinates
  st_transform(crs = 32748) # IN projected coordinate system
```

```{r}
kable(head(fire_sf,5))
```

```{r}
summary(fire_sf)
```

Because **ppp** object only accept numerical or character as mark. The code chunk below is used to convert data type of *acq_date* to numeric:

```{r}
fire_sf <- fire_sf %>%
  mutate(DayofYear = yday(acq_date),
         Month_num = month(acq_date),
         Month_fac = month(acq_date, 
                           label = TRUE, 
                           abbr = FALSE))
```

# 5 Visualize the Fire Points

## 5.1 Overall Plot

```{r}
tm_shape(kbb_sf)+
  tm_polygons()+
  tm_shape(fire_sf)+
  tm_dots(size = 0.2, shape = 16)+
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6")
  
```

## 5.2 Forest Fires Distribution by Month

```{r}
tm_shape(kbb_sf)+
  tm_polygons()+
  tm_shape(fire_sf)+
  tm_dots(size = 0.2, shape = 16)+
  tm_title("KKB Monthly Distribution of Forest Fires in 2023",
           position = tm_pos_out(cell.h = "center",
                                 cell.v = "top",
                                 pos.h = "center"))+ 
  tm_facets(nrow = 3, ncol = 4,
            by = "Month_fac",
            free.coords = FALSE)+ 
  tm_layout(frame = FALSE,
            bg.color = "#f6f6f6",
            panel.label.bg.color = "#AFCD6F",
            inner.margins = c(0, 0.05, 0.2, 0.05)) #bottom, left, top, right
```

As shown in the graph above, KBB overall experienced fewer forest fires from December to June, while more incidents occurred from July to November.

Spatially, more frequent and densely clustered incidents occurred in the northern region during July and August, whereas the southern region saw increased fire activity from September to November.

::: callout-tip
## Main Title of the Graph

According to the tmap documentation, the `main.title` setting within `tm_layout()` is deprecated, and it now recommends using `tm_title()` instead.

To add a main title for the entire graph, use `tm_title(position = tm_pos_out())`. For a top-centered title, adjust the position parameters as follows:

``` r
tm_title("Title", 
         position = tm_pos_out(cell.h = "center", 
                              cell.v = "top", 
                              pos.h = "center"))
```

or

``` r
tm_title("Title", 
         position = tm_pos_out("center","top","center")) 
```
:::

::: callout-tip
-   **`free.coords = TRUE`** shows maximum detail by allowing each facet to zoom independently.
-   **`free.coords = FALSE`** maintains spatial reference points across all facets.
:::

# 6 Compute STKDE by Month

In the previous hands-on exercise, we learned to use spatstat's `density()` function to plot KDE maps. The more advanced spatio-temporal KDE function (`spattemp.density()` from **sparr**) can analyze not only spatial data but also incorporates temporal data.

In this section, we will learn how to compute STKDE by using [`spattemp.density()`](https://tilmandavies.github.io/sparr/reference/spattemp.density.html) of **sparr** package.

The important arguments of the function includes:

| Argument | Description |
|------------------|------------------------------------------------------|
| **h (spatial)** | Fixed bandwidth to smooth the spatial margin. A numeric value \> 0. If unsupplied, the oversmoothing bandwidth is used as per [`OS`](https://tilmandavies.github.io/sparr/reference/OS.html). |
| **lambda (temporal)** | Fixed bandwidth to smooth the temporal margin; a numeric value \> 0. If unsupplied, the function internally computes the Sheather-Jones bandwith using [`bw.SJ`](https://rdrr.io/r/stats/bandwidth.html) (Sheather & Jones, 1991). |
| **tt (attribute)** | A numeric vector of equal length to the number of points in `pp`, giving the time corresponding to each spatial observation. If unsupplied, the function attempts to use the values in the [`marks`](https://rdrr.io/pkg/spatstat.geom/man/marks.html) attribute of the [`ppp.object`](https://rdrr.io/pkg/spatstat.geom/man/ppp.html) in `pp`. |

## 6.1 Extract forest fires by month

The code chunk below is used to keep only the month info and geometry of *fire_sf* sf data.frame.

The result, `fire_month`, is still an `sf` object with geometry and one attribute column (month):

```{r}
fire_month <- fire_sf %>%
  select(Month_num)
kable(head(fire_month,5))
```

## 6.2 Create ppp Object

Once our data is in `ppp` format, we can use `spatstat` functions like: density.ppp(), Kest(),envelope() and `sparr` function: spattemp.density()

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

Next, we use summary() to check the points

```{r}
summary(fire_month_ppp)
```

Next, we check if there are any duplicated point events:

```{r}
any(duplicated(fire_month_ppp))
```

## 6.3 Create the Study Window

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

As a good practice, `plot()` is used to plot fire_month_owin, so that we can examine the correctness of the output object:

```{r}
par(bg = "#f6f6f6")

plot(fire_month_owin,
     main = "Fire Incidents by Month")
```

## 6.4 Compute Spatio-temperal KDE

`spattemp.density()` of sparr package is used to compute the STKDE:

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

## 6.5 Plot the Spatio-temperal KDE Object

`plot()` of R base is used to the KDE for between July 2023 - December 2023:

```{r}
tims <- c(7,8,9,10,11,12)
par(bg = "#f6f6f6",mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}
```

# 7 Compute Spatio-temperal KDE by Day of Year

## 7.1 Create ppp

```{r}
fire_yday_ppp <- fire_sf %>% 
  select(DayofYear) %>%
  as.ppp()
summary(fire_yday_ppp)
```

## 7.2 Create the Study Window

```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

## 7.3 Compute Spatio-temperal KDE

```{r}
kde_yday <- spattemp.density(
  fire_yday_owin)
summary(kde_yday)
```

## 7.4 Plot the Spatio-temperal KDE Object

For efficiency purposes, we focus on October since it had the highest density of forest fires.

First, we filter data to include only October incidents:

```{r}
select_yday <- fire_sf %>%
  filter(Month_fac == "October") %>%
  pull(DayofYear) %>%
  unique()
select_yday
```

Next, we use the same for-loop structure to generate daily plots for October:

```{r}
par(bg = "#f6f6f6", mfcol=c(2,3))  # 2x3 plots per page
for (i in select_yday) {
  plot(kde_yday, i,
       override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at Day of Year", i))

}

```

# 8 Compute STKDE by Day of Year (Improved method)

One of the nice function provides in sparr package is `BOOT.spattemp()`. It support bandwidth selection for standalone spatio-temporal density/intensity based on bootstrap estimation of the MISE, providing an isotropic scalar spatial bandwidth and a scalar temporal bandwidth.

Code chunk below uses `BOOT.spattemp()` to determine both the spatial bandwidth and the scalar temporal bandwidth.

```{r}
set.seed(1234)
BOOT.spattemp(fire_yday_owin) 
```

## 8.1 Compute Spatio-temperal KDE

Now, the STKDE will be derived by using h and lambda values derive in previous step

```{r}
kde_yday <- spattemp.density(fire_yday_owin,
                             h = 9105.73611,
                             lambda = 18.85762)
summary(kde_yday)
```

## 8.2 Plot the Spatio-temperal KDE Object

```{r}
par(bg = "#f6f6f6", mfcol=c(2,3))  # 2x3 plots per page
for (i in select_yday) {
  plot(kde_yday, i,
       override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at Day of Year", i))

}
```

Compared to the orignal result, the `boot.spattemp()` suggests fire data benefits from **finer spatial resolution** but **broader temporal smoothing** for better density estimation:

|                   | w/o BOOT.spattemp | with BOOT.spattemp |
|-------------------|-------------------|--------------------|
| h (spatial)       | 15,102.47         | 9105.736           |
| lambda (temporal) | 6.32              | 18.8576            |

Compared to the original STKDE graphs, we can now clearly see the hot spots in the southern region during October.

# 9 Reference

-   Kam, T.S. (2025). [Spatio-Temporal Point Patterns Analysis](https://r4gdsa.netlify.app/chap06#computing-spatio-temporal-kde-1)
