---
title: "Hands-on 2.b: 2nd Order Spatial Point Patterns Analysis Methods"
author: Johsuan Huang
date: 2025-09-03
date-modified: "last-modified"
description: ""
format: html
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

Second-order spatial point pattern analysis examines the spatial relationships between points in a pattern, specifically focusing on how the presence of one point influences the location of others (eg. 7-11 and Family Marts always open near to each other for the traffics). It goes beyond simply describing the overall density of points (first-order effects) by investigating clustering, dispersion, or randomness at various spatial scales.

Using appropriate functions of [spatstat](https://cran.r-project.org/web/packages/spatstat/), this hands-on exercise aims to discover the spatial point processes of childecare centres in Singapore.

The specific questions we would like to answer are as follows:

-   are the childcare centres in Singapore randomly distributed throughout the country?

-   if the answer is not, then the next logical question is where are the locations with higher concentration of childcare centres?

# 2 The Data

The data we used is the same as Hands-on Ex2-1:

-   [*Child Care Services*](https://data.gov.sg/datasets?query=Child+Care+Services&resultId=d_2bbcc5e944a9f2d7d40c78097ab652fc&page=1) data from data.gov.sg, a point feature data providing both location and attribute information of childcare centres.
-   [*Master Plan 2019 Subzone Boundary (No Sea)*](https://data.gov.sg/datasets?query=Master+Plan+2019+Subzone+Boundary+(No+Sea)&page=1&resultId=d_0900fdcfdb4666fe2a630f37a532fc06), a polygon feature data providing information of URA 2019 Master Plan Planning Subzone boundary data.

# 3 Import Libraries

+---------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Library                                                             | Desc                                                                                                                                                                                                                                                    |
+=====================================================================+=========================================================================================================================================================================================================================================================+
| [**sf**](https://r-spatial.github.io/sf/)                           | handling geospatial data                                                                                                                                                                                                                                |
+---------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| [**tmap**](https://cran.r-project.org/web/packages/tmap/index.html) | visualization for geospatial data                                                                                                                                                                                                                       |
+---------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| tidyverse                                                           | handling data including packages **readr**, **tidyr and dplyr**                                                                                                                                                                                         |
+---------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| [**rvest**](https://rvest.tidyverse.org/)                           | scraping (or harvesting) data from html and web pages                                                                                                                                                                                                   |
+---------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| [**spatstat**](https://spatstat.org/)                               | for point pattern analysis                                                                                                                                                                                                                              |
+---------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| [**terra**](https://rspatial.github.io/terra/index.html)            | a modern spatial data analysis package to replace the raster package. terra provides functionalities for creating, reading, manipulating, and writing raster and vector data, and it’s built on top of GDAL and PROJ libraries for enhanced performance |
+---------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
pacman::p_load(sf, terra, spatstat,tmap, rvest, tidyverse)
```

# 4 Import Data

```{r}
mpsz_cl <- read_rds("Data/geospatial/mpsz_cl.rds")

childcare_sf <- st_read("Data/geospatial/ChildCareServices.kml")%>%
  st_zm(drop = TRUE, what='ZM') %>% 
  st_transform(crs = 3414)

st_geometry(childcare_sf)

```

```{r}
childcare_ppp <- as.ppp(childcare_sf) %>%
          rjitter(retry = TRUE,
          nsim = 1,
          drop = TRUE) #disperse overlapping points
```

```{r}
pg <- mpsz_cl %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_cl %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_cl %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_cl %>%
  filter(PLN_AREA_N == "JURONG WEST")
```

```{r}
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)
```

```{r}
childcare_pg_ppp = childcare_ppp[pg_owin]
childcare_tm_ppp = childcare_ppp[tm_owin]
childcare_ck_ppp = childcare_ppp[ck_owin]
childcare_jw_ppp = childcare_ppp[jw_owin]
```

```{r}
childcare_pg_ppp_km = rescale.ppp(childcare_pg_ppp,1000,"km")
childcare_tm_ppp_km = rescale.ppp(childcare_tm_ppp,1000,"km")
childcare_ck_ppp_km = rescale.ppp(childcare_ck_ppp,1000,"km")
childcare_jw_ppp_km = rescale.ppp(childcare_jw_ppp,1000,"km")
```

```{r}
par(bg="#f6f6f6",mfrow=c(2,2))
plot(unmark(childcare_pg_ppp_km), 
  main="Punggol")
plot(unmark(childcare_tm_ppp_km), 
  main="Tampines")
plot(unmark(childcare_ck_ppp_km), 
  main="Choa Chu Kang")
plot(unmark(childcare_jw_ppp_km), 
  main="Jurong West")
```

# 5 Second-order Spatial Point Patterns Analysis

In this exercise, we will use four functions to analyze the spatial point patterns:

+---------------------------+------------+-----------------------------------------------------------------+-----------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+
| Function                  | Focus Type | What It Measures                                                | Analogy / Example                                                                 | Remark                                                                                     |
+===========================+============+=================================================================+===================================================================================+============================================================================================+
| `Gest()`                  | **Local**  | Distance from a point to its **nearest neighbor**               | For each childcare centre, how far is the nearest other centre?                   | -   If most centres are **very close** to another → **clustering**                         |
|                           |            |                                                                 |                                                                                   | -   If most centres are **far apart** → **dispersion**                                     |
| Nearest Neighbor Distance |            |                                                                 |                                                                                   |                                                                                            |
+---------------------------+------------+-----------------------------------------------------------------+-----------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+
| `Fest()`                  | **Local**  | Distance from **random spot** to nearest point                  | From random locations in the city, how far is it to the nearest childcare centre? | -   If the city is well-covered, **most locations are close** to a centre → **clustering** |
|                           |            |                                                                 |                                                                                   |                                                                                            |
| Empty Space Function      |            |                                                                 |                                                                                   | -   If large areas are far from a centre → **dispersion**                                  |
+---------------------------+------------+-----------------------------------------------------------------+-----------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+
| `Kest()`                  | **Global** | Counts how many other points are within distance `r` of a point | For every centre, how many others fall within a circle of radius `r`?             | Compares that number to what you'd expect under **random** placement                       |
|                           |            |                                                                 |                                                                                   |                                                                                            |
| Ripley's K Function       |            |                                                                 |                                                                                   |                                                                                            |
+---------------------------+------------+-----------------------------------------------------------------+-----------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+
| `Lest()`                  | **Global** | Same as K, but easier to interpret                              | Like K, but transformed so a straight line (accumulation) means "no pattern"      | -   If line **stays above** diagonal → clustering                                          |
|                           |            |                                                                 |                                                                                   |                                                                                            |
| L Function                |            |                                                                 |                                                                                   | -   If line **stays below** → regular spacing                                              |
+---------------------------+------------+-----------------------------------------------------------------+-----------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+

## 5.1 Analysing Spatial Point Process Using G-Function

The G function measures the distribution of the distances from an arbitrary event to its nearest event. In this section, you will learn how to compute G-function estimation by using [*Gest()*](https://rdrr.io/cran/spatstat/man/Gest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using [*envelope()*](https://rdrr.io/cran/spatstat/man/envelope.html) of **spatstat** package

### 5.1.1 Choa Chu Kang planning area

#### 5.1.1.1 Computing G-function estimation

```{r}
set.seed(1234)
```

The code chunk below is used to compute G-function using *Gest()* of **spatat** package.

```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
par(bg = "#f6f6f6")
plot(G_CK, xlim=c(0,500))
```

Since the observed G-function rises **faster** than expected under CSR, it means that points are **closer together than expected**, suggesting a **clustered** spatial pattern. However, we still need further test to conclude if it's clustered.

#### 5.1.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.
-   H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.
-   The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with G-fucntion:

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

```{r}
par(bg = "#f6f6f6")
plot(G_CK.csr)
```

### 5.1.2 Tampines planning area

#### 5.1.2.1 Computing G-function estimation

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
par(bg = "#f6f6f6")
plot(G_tm)
```

#### 5.1.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Tampines are randomly distributed.
-   H1= The distribution of childcare services at Tampines are not randomly distributed.
-   The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing:

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

```{r}
par(bg = "#f6f6f6")
plot(G_tm.csr)
```

## 5.2 Analysing Spatial Point Process Using F-Function

The F function estimates the empty space function F(r) or its hazard rate h(r) from a point pattern in a window of arbitrary shape. In this section, you will learn how to compute F-function estimation by using [*Fest()*](https://rdrr.io/cran/spatstat/man/Fest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using [*envelope()*](https://rdrr.io/cran/spatstat/man/envelope.html) of **spatstat** package.

### 5.2.1 Choa Chu Kang planning area

#### 5.2.1.1 Computing F-function estimation

The code chunk below is used to compute F-function using *Fest()* of **spatat** package.

```{r}
F_CK = Fest(childcare_ck_ppp)
par(bg = "#f6f6f6")
plot(F_CK)
```

#### 5.2.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.
-   H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.
-   The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with F-fucntion:

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r}
par(bg = "#f6f6f6")
plot(F_CK.csr)
```

### 5.2.2 Tampines planning area

#### 5.2.2.1 Computing F-function estimation

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
par(bg = "#f6f6f6")
plot(F_tm)
```

#### 5.2.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Tampines are randomly distributed.
-   H1= The distribution of childcare services at Tampines are not randomly distributed.
-   The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing:

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

```{r}
par(bg = "#f6f6f6")
plot(F_tm.csr)
```

## 5.3 Analysing Spatial Point Process Using K-Function

The K-function measures the number of events found up to a given distance of any particular event. In this section, you will learn how to compute K-function estimates by using [*Kest()*](https://rdrr.io/cran/spatstat/man/Kest.html) of **spatstat**package. You will also learn how to perform monta carlo simulation test using *envelope()* of spatstat package.

### 5.3.1 Choa Chu Kang planning area

#### 5.3.1.1 Computing K-function estimation

The code chunk below is used to compute K-function using *Kest()* of **spatat** package.

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
par(bg = "#f6f6f6")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

#### 5.3.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.
-   H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.
-   The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with F-fucntion:

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
par(bg = "#f6f6f6")
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### 5.3.2 Tampines planning area

#### 5.3.2.1 Computing K-function estimation

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
par(bg = "#f6f6f6")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### 5.3.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Tampines are randomly distributed.
-   H1= The distribution of childcare services at Tampines are not randomly distributed.
-   The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing:

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
par(bg = "#f6f6f6")
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

## 5.4 Analysing Spatial Point Process Using L-Function

In this section, we will learn how to compute L-function estimation by using [*Lest()*](https://rdrr.io/cran/spatstat/man/Lest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using *envelope()* of spatstat package.

### 5.4.1 Choa Chu Kang planning area

#### 5.4.1.1 Computing L-function estimation

The code chunk below is used to compute K-function using *Lest()* of **spatat** package.

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")

par(bg = "#f6f6f6")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

#### 5.4.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.
-   H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.
-   The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with F-fucntion:

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
par(bg = "#f6f6f6")
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

### 5.4.2 Tampines planning area

#### 5.4.2.1 Computing L-function estimation

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
par(bg = "#f6f6f6")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### 5.4.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

-   Ho = The distribution of childcare services at Tampines are randomly distributed.
-   H1= The distribution of childcare services at Tampines are not randomly distributed.
-   The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing:

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
par(bg = "#f6f6f6")
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```

Tampines form clusters around 50m. However, CK forms cluster

# 6 Reference

-   Kam, T.S. (2025). [2nd Order Spatial Point Patterns Analysis Methods](https://r4gdsa.netlify.app/chap05#analysing-spatial-point-process-using-k-function)
