---
title: "In-class Exercise 4: GWmodel"
author: Huang Johsuan
date: "September 20, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1 Import Libraries

```{r}
pacman::p_load(sf, ggstatsplot, 
               tmap,tidyverse,knitr,GWmodel)
```

# 2 Import Data

```{r}
#| echo: false
hunan <- st_read(dsn="data/geospatial",
               layer = "Hunan")%>%
  st_transform(crs=4490)
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")

```

```{r}
hunan_sf <- left_join(hunan,hunan2012,
                   by = "County") %>%
  select(1:3,7,15,16,31,32)
kable(head(hunan_sf,5))
```

# 3 Convert sf to sp format

GWmodel is an older R package that requires spatial data to be in the 'sp' format, rather than the newer 'sf' format:

```{r}
hunan_sp <- hunan_sf %>%
  as_Spatial()
hunan_sp
```

# 4 Compute Fixed Distance Using GWmodel

There are two methods provided by GWmodel: cross validation and AIC. Note the default return value is in kilometer.

## 4.1 CV method

```{r}
bw_CV <- bw.gwr(GDPPC~1,
                data = hunan_sp,
                approach = "CV", # cross validation method
                adaptive = FALSE,
                kernel = "bisquare",
                longlat = T # True, indicate the that data is in long, lat
                ) 
bw_CV
```

The algorithm will go through several iterations and then converge to the optimal bandwith: 76.29126

## 4.2 AIC method

```{r}
bw_AIC <- bw.gwr(GDPPC~1,
                data = hunan_sp,
                approach = "AIC", 
                adaptive = FALSE,
                kernel = "bisquare",
                longlat = T # True, indicate the that data is in long, lat
                ) 
bw_AIC
```

# 5 Summary Statistics

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               adaptive = FALSE,
               longlat = T
               )
gwstat
```

## 5.1 Transform the output into df format

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
kable(head(gwstat_df))
```

## 5.2 Combine the statistics result

Since `gwstat_df` does not contain a primary key and both datasets share the same row order, we use `cbind()` instead of a join

```{r}
hunan_gstat <- cbind(hunan_sf,gwstat_df)
kable(head(hunan_gstat))
```

```{r}
#| fig.width: 12
#| fig.height: 10

LM <- tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LM") +
  tm_text("NAME_3", size=0.6)+
 tm_title("GDPPC Local Means",
           position = tm_pos_out("center", "top", "center")) +
  tm_layout(
    legend.position = c("left", "bottom"),
    bg.color = "#f6f6f6",
    frame = FALSE)

LSD <- tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LSD") +
  tm_text("NAME_3", size=0.6)+
 tm_title("GDPPC Local Standard Deviation",
           position = tm_pos_out("center", "top", "center")) +
  tm_layout(
    legend.position = c("left", "bottom"),
    bg.color = "#f6f6f6",
    frame = FALSE)

LVar <- tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LVar") +
  tm_text("NAME_3", size=0.6)+
 tm_title("GDPPC Local Variance",
           position = tm_pos_out("center", "top", "center")) +
  tm_layout(
    legend.position = c("left", "bottom"),
    bg.color = "#f6f6f6",
    frame = FALSE)


tmap_arrange(LM, LSD,LVar, asp=1, nrow =2, ncol=2)
```
